---
---
title: "FRAX calculator for patients without fracture pulled by month"
author: "Anthony Criscitiello and Ellen Quillen"
date: "9/23/2021"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "FRAX calculator for I2b2 results"
output: html_document
---
```{r setup}
library(tidyverse)
library(readxl)
library(readr)
library(lubridate)
library(tibble)
library(purrr)
library(stringr)
```
^I find it helpful to set everything up at the beggining and you can also make some modifications to how this Rmd turns into an html. Here is cheatsheet on options for R markdown: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
##Introduction
```{r load FRAX ppr charts and patient data}
#load FRAX paper charts (10-year risk of major osteoporotic fracture based on BMI)
FRAX <-  read_csv("H:/FRAX.csv", col_types = "dccdddddd")
FRAX <- FRAX[-c(1)]
names(FRAX)<- c("GENDER", "RACE", "AGE","CRF","BMI","Risk", "Range_lower", "Range_upper")
#Note that unlike prior data, one excel file contains all information for all patients seen in the ED over the course of one month
tibble_names <- ("H:/Health Control (Hip fx)/2015_acriscit.xlsx")

#Function to import excel data with multiple tabs as a list
read_excel_allsheets <- function(filename, tibble = TRUE) {
  tab_names <- excel_sheets(path = filename)
#Below is the slow line
filename <- map(tab_names, ~read_excel(path = filename, sheet = .x))
filename <- filename %>% set_names(tab_names)
#Next Remove patients and visits because they do not fit the common column naming convention below
#filename_patients <- filename$Patients
#filename_visits <- filename$Visits
#filename <- within(filename, rm(Patients, Visits))
}
filename <- tibble_names
tibble_names <- read_excel_allsheets(filename)
#Patients <- filename_patients
#tibble_names_visits <- filename_visits

Patients <- tibble_names$Patients
Visits <- tibble_names$Visits
tibble_names <- within(tibble_names, rm(Patients, Visits))

#Break the tibble down into components rather than conglomerate into a large tibble. This will significantly speed up the filter function used to search for specific data elements later in this code
#put this into the function above
tibble_titles <- c(
  "Procedures",
  "Diagnoses",
  "Labs",
  "Meds",
  "Vitals", 
  "Smoking"
  )
 test2 <- list()
tab_names <- excel_sheets(path = "H:/Health Control (Hip fx)/2013_acriscit.xlsx")
#The below lines combines tabs of the same type, for example all the Meds tabs or all the Labs tabs and forms a new list
for (i in 1:length(tibble_titles)) {
 tab_names_x <- tab_names %>% as_tibble() %>% filter(grepl(tibble_titles[i], value))
 x <- tibble_names[tab_names_x$value] %>% bind_rows()
 test2 <- append(test2, list(x))
}
#Be sure that the order in tibble_titles actually corresponds to the info in test2
names(test2) <- tibble_titles
#replace tibble names
tibble_names <- test2

rm(test2, x)

#Rename all the columns in each list element of tibble_names and change DATE to a date format
col_names <- c(
  "ID", 
  "VISIT", 
  "DATE",
  "CODE",
  "DESC", 
  "TEXT_VALUE", 
  "NUMERIC_VALUE", 
  "UNITS",
  "MODIFIER"
)
for (i in 1:length(tibble_names)){
  names(tibble_names[[i]]) <- col_names[1:ncol(tibble_names[[i]])]
   tibble_names[[i]]$DATE <- mdy_hm(
    tibble_names[[i]]$DATE, 
    quiet = TRUE
    )
}

#Rename the columns in 'Patients'
col_names <- c(
  "ID", 
  "AGE_NOW", 
  "AGE_AT_DEATH",
  "BIRTH_DATE",
  "DEATH_DATE",
  "MATCH",
  "GENDER", 
  "RACE", 
  "MIXED RACE"
)
names(Patients) <- col_names[1:ncol(Patients)]
#Make DATE column into date format
  Patients <- Patients %>%
    mutate(BIRTH_DATE = mdy_hm(BIRTH_DATE) %>% as.Date()) %>%
    mutate(DEATH_DATE = mdy_hm(DEATH_DATE) %>% as.Date())

#Does the same for visits
  col_names <- c(
  "ID", 
  "VISIT", 
  "START_DATE",
  "END_DATE",
  "AGE_AT_VISIT",
  "LOS",
  "ENCOUNTER_TYPE",
  "CCI"
)
#Applies the new column names to each list element in 'tibble_names' and converts dates to mdy_hms
  names(Visits) <- col_names[1:ncol(Visits)]
  
  #Convert visits dates to dates and use as.Date to remove the time element
  Visits <- Visits %>% mutate(START_DATE = as.Date(mdy_hm(START_DATE)))
  Visits <- Visits %>% mutate(END_DATE = as.Date(mdy_hm(END_DATE)))

  #One visit is often tracked as multiple visits on the same date. Remove duplicate visits
  Visits <- Visits %>%
    distinct(ID, START_DATE, .keep_all = TRUE) %>% group_by(ID) %>% arrange(START_DATE, .by_group = TRUE)
```
If files are in sepearate docs use this to load

FRAX <-  read_csv("H:/Hip fracture/FRAX.csv", col_types = "dccdddddd")
FRAX <- FRAX[-c(1)]
names(FRAX)<- c("GENDER", "RACE", "AGE","CRF","BMI","Risk", "Range_lower", "Range_upper")

Diagnoses <- ("H:/raw data/Health Control (Hip fx)/2014/Diagnoses_acriscit.xlsx")
History <- ("H:/raw data/Health Control (Hip fx)/2014/Smoking_acriscit.xlsx")
Labs <- ("H:/raw data/Health Control (Hip fx)/2014/Labs_acriscit.xlsx")
Meds <- ("H:/raw data/Health Control (Hip fx)/2014/Meds_acriscit.xlsx")
Patients <- ("H:/raw data/Health Control (Hip fx)/2014/Patients_acriscit.xlsx")
Procedures <- ("H:/raw data/Health Control (Hip fx)/2014/Procedures_acriscit.xlsx")
Vitals <- ("H:/raw data/Health Control (Hip fx)/2014/Vitals_acriscit.xlsx")

#Function to import excel data with multiple tabs as a list
read_excel_allsheets <- function(filename, tibble = TRUE) {
  tab_names <- excel_sheets(path = filename)
  #Below is the slow line
  filename <- map(tab_names, ~read_excel(path = filename, sheet = .x, col_types = "text"))
  filename <- filename %>% set_names(tab_names) %>% bind_rows ()
}

#The function above will convert each file below into a list that is easy to work with
filename <- Diagnoses
Diagnosis <- read_excel_allsheets(filename)

filename <- History
History <- read_excel_allsheets(filename)

filename <- Meds
Meds <- read_excel_allsheets(filename)

filename <- Patients
Patients <- read_excel_allsheets(filename)

filename <- Procedures
Procedures <- read_excel_allsheets(filename)

filename <- Vitals
Vitals <- read_excel_allsheets(filename)
  
filename <- Labs
Lab <- read_excel_allsheets(filename)


#In this chunk we are organizing data so that relevant information can be more easily collected
#- All the documents we imported above will be combined into one large tibble
#- We will create an empty tibble that will be filled with info to calculate FRAX score at the time of fracture for each patient

#Accumulate data tibbles as a list
tibble_names <- list(
  Vitals, 
  Meds, 
  Lab, 
  Procedures, 
  Diagnosis,
  History
  )
#Title each list element appropriately
tibble_titles <- c(
  "Vitals", 
  "Meds", 
  "Lab", 
  "Procedures",
  "Diagnosis",
  "History"
  )
names(tibble_names) <- tibble_titles

#rename columns and remove spaces from column names
col_names <- c(
  "ID", 
  "VISIT", 
  "DATE",
  "CODE",
  "DESC", 
  "TEXT_VALUE", 
  "NUMERIC_VALUE", 
  "UNITS",
  "MODIFIER"
)
#This loop applies the new column names to each list element in 'tibble_names' and converts dates to mdy_hms
for (i in 1:length(tibble_names)){
  names(tibble_names[[i]]) <- col_names[1:ncol(tibble_names[[i]])]
  tibble_names[[i]]$DATE <- mdy_hm(
    tibble_names[[i]]$DATE, 
    quiet = TRUE
    )
}

#remove the xl files. Note that if you want to search one of these files rather than x (because it is faster) you can search tibble_names$...
rm(Diagnoses, Labs, Meds, Vitals, History)

#Do the same process for demographic data
col_names <- c(
  "ID", 
  "AGE_NOW", 
  "AGE_AT_DEATH",
  "BIRTH_DATE",
  "DEATH_DATE",
  "MATCH",
  "GENDER", 
  "RACE", 
  "MIXED RACE", 
  "ETHNICITY"
)
names(Patients) <- col_names[1:ncol(Patients)]

  Patients <- Patients %>%
    mutate(BIRTH_DATE = mdy_hm(BIRTH_DATE))
  
 Patients <- Patients %>%
    mutate(DEATH_DATE = mdy_hm(DEATH_DATE))
 
 #Do the same for Visits
 col_names <- c(
  "ID", 
  "", 
  "AGE_AT_DEATH",
  "BIRTH_DATE",
  "DEATH_DATE",
  "MATCH",
  "GENDER", 
  "RACE", 
  "MIXED RACE", 
  "ETHNICITY"
)
names(Patients) <- col_names[1:ncol(Patients)]

  Patients <- Patients %>%
    mutate(BIRTH_DATE = mdy_hm(BIRTH_DATE))
  
 Patients <- Patients %>%
    mutate(DEATH_DATE = mdy_hm(DEATH_DATE))
```
Create a tibble with a row for each patient with the date of their ED visit and Diagnoses at visit
```{r ED}
date_lower <- mdy("5/1/2013")
date_upper <- mdy("5/31/2014")
ID_EM <- tibble_names$Diagnoses %>% filter(TEXT_VALUE == "Primary" & DATE > date_lower & DATE < date_upper) %>% group_by('ID') %>% arrange(DATE, .group_by = TRUE) %>% distinct(ID, .keep_all = TRUE)
```
#The below chunk will create a tibble for each FRAX risk factor that includes all patients. For example 'Smoker' is a tibble that contains smoking information on all patients with a history of smoking. An individual patient will be selected from this tibble within the loop. This was done to increase the efficiency of the loop by minimizing the number of functions within the loop
```{r risk_factors}
# alternative to 'filter(grepl())' is library(stringr), x %>% filter(str_detect(rowname, "^ICD-10 code")). '^' means starts with
Fracture <- tibble_names$Diagnoses %>% 
     filter(grepl('[Ff]racture', DESC)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)
#Smoker at the time of the fracture
Smoker <- tibble_names$Smoking %>% 
     filter(grepl('[Ss]moker', DESC) & grepl('[Cc]urrent', DESC)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)
#A Diagnoses of alcohol use (ICD-10 codes containing F10 or corresponding ICD-9 codes (303 alcohol dependance, 291 alcohol induced mental disorder, 3050 alcohol abuse, K70 alcoholic liver disease), which encompasses alcohol abuse, alcohol dependence and alcohol use, unspecified) is used as a surrogate for "3 or more units of alcohol per day" as specified by FRAX
EToH <- tibble_names$Diagnoses %>% 
     filter(grepl('F10|3050|291|303|V11.3|K70', CODE)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)
#RA Diagnoses of Rheumatoid Arthritis 
RA <- tibble_names$Diagnoses %>% 
     filter(grepl('M05|M06|714', CODE)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)
#According to FRAX this includes osteoporosis (M81, 7330) and diseases strongly associated with osteoporosis including type I (insulin dependent) diabetes (E10, 250), osteogenesis imperfecta in adults (Q78.0, 75651), untreated long-standing hyperthyroidism E05, 242)(these are the codes for hyperthyroidism irrespective of treatment), hypogonadism (E29.1, 2572 (male), E28.3, 25631, 25639 (female) including) premature menopause (<45 years), chronic malnutrition (intestinal malabsorption K90, 579, nutritional malabsorption E41-46, 261-263, 579), or malabsorption and chronic liver disease (571-573, K71-K76)
#An alternative to malabsorption could be to consider BMI <15
Osteoporosis <- tibble_names$Diagnoses %>% 
     filter(grepl('M81|7330|E10|250|Q78.0|75651|E05|242|E29.1|2572|E28.3|25631|25639|k90|579|E41|E42|E43|E44|E45|E46|261|262|263|579|571|572|573|K71|K72|K73|K74|K75|K76', CODE)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)
#FRAX considers parental hip fracture a risk. We have substituted parental osteoporosis for this risk factor because there is a searchable diagnostic code
Parent_Osteoporosis <- tibble_names$Diagnoses %>% 
     filter(grepl('Z82.62|V17.81', CODE)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)
#Current exposure to oral glucocorticoids or exposure to oral glucocorticoids for more than 3 months at a dose eqivalent to prednisolone > 5mg daily
Glucocorticoid <- tibble_names$Meds %>% 
     filter(MODIFIER == 'mg' &
              grepl('[Bb]eclomethasone|[Bb]etamethasone|[Bb]udesonide|[Cc]ortisone|[Dd]examethasone|[Hh]ydrocortisone|[Mm]ethylprednisolone|[Pp]rednisolone|[Pp]rednisone|[Tt]riamcinolone', DESC)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)

```
The below loop will collect and distill information relevant to calculating FRAX score FOR PATIENTS WHO HAVE HAD A FRACTURE ONLY. Each cycle of this loop will collect data for 1 patient and will populate the tibble 'EM_Dates'
```{r loop}
#Create an empty tibble ('EM_Dates') with the below columns. It will later be populated by each cycle of the loop
EM_Dates <- tibble(
  ID = numeric(),
  DATE = mdy_hms(),
  CODE = character(),
  DESC = character(),
  HT.in = numeric(), 
  WT.in = numeric(),
  Birth_Date = mdy_hms(),
  ageatEM = numeric(),
  ageatdeath = numeric(),
  GENDER = character(), 
  Og_RACE = character(), 
  CRF = numeric(),
  #CRF is Combined Risk Factors
  BMI = numeric()
)
#Tibble where the risk factors will be compiled
CRF <- tibble(
  ID = numeric(),
  Fracture = character(),
  Smoker = character(),
  EToH = character(),
  RA = character(),
  Osteoporosis = character(),
  Parent_Osteoporosis = character(),
  Glucocorticoid_rf = character()
)
#Collect risk factors into above tibble. If there is no observation prior to DATE, it is assumed that the risk factor was not present. For example, if there is no prior fracture documented, it is assumed that there was no prior fracture

for (i in 1:nrow(ID_EM)){
  id <- as.double(ID_EM$ID[i])
loop_ID <- ID_EM %>% 
    filter(ID == id)
#Below are common variables that will be used throughout the loop
date <- loop_ID$DATE[1]
CODE <- loop_ID$CODE[1]
DESC <- loop_ID$DESC[1]
#REMOVE THIS IF YOU ARE NOT LOOKING FOR HIP FRACTURES ONLY!!!
#####
#Create a tibble with one patients into
loop_ID <- tibble_names$Vitals %>% 
    filter(ID == id)
#Select only the info prior to fracture
loop_ID_prefrac <- loop_ID %>% filter(DATE <= date)
#Height and weight closest to the time of fracture
HT <- (loop_ID_prefrac %>% 
         filter(CODE == "HT") %>% 
         arrange(desc(DATE)) %>% 
         select(NUMERIC_VALUE)
       )[1,1]
WT <- (loop_ID_prefrac %>% 
         filter(CODE == "WT") %>% 
         arrange(desc(DATE)) %>% 
         select(NUMERIC_VALUE)
       )[1,1]
#If there is no height or weight prior to the fracture date, use the oldest measurement
HT <- as.double(ifelse(is.na(HT), ((
  loop_ID %>% 
    filter(CODE == "HT") %>% 
    arrange(desc(DATE)) %>% 
    select(NUMERIC_VALUE))[1,1]),
  HT))
WT <- as.double(ifelse (is.na(WT), ((
  loop_ID %>% 
    filter(CODE == "WT") %>% 
    arrange(desc(DATE)) %>% 
    select(NUMERIC_VALUE))[1,1]), 
  WT))
  
rm(loop_ID, loop_ID_prefrac)
#Calculate BMI
BMI <- (703*as.double(WT)/as.double(HT)^2)

yID <- Patients %>% filter(ID == id)
#Calcuate age at time of fracture
Birth_Date <- (yID %>% select(BIRTH_DATE))[[1]]
ageatEM <- (date - Birth_Date) %>% as.double/365.25
ageatdeath <- as.double(yID %>% select(AGE_AT_DEATH))
#Select gender
Gender <- yID %>% select(GENDER)
#Select race
Race <- yID %>% select(RACE)
rm(yID)
# alternative to 'filter(grepl())' is library(stringr), x %>% filter(str_detect(rowname, "^ICD-10 code")). '^' means starts with
Fracture_loop <- as.character(
  (filter(Fracture, id == ID & DATE <= date) %>%
  select(DESC))[1,2]
  )
#Smoking within the last 2 years
Smoker_loop <- as.character(
  (filter(Smoker, id == ID & DATE <= date & DATE >= date-years(2)) %>% 
     select(DESC))[1,2]
  )
#A Diagnoses of alcohol use (ICD-10 codes containing F10 or corresponding ICD-9 codes (303 alcohol dependance, 291 alcohol induced mental disorder, 3050 alcohol abuse), which encompasses alcohol abuse, alcohol dependence and alcohol use, unspecified) is used as a surrogate for "3 or more units of alcohol per day" as specified by FRAX
EToH_loop <- as.character(
  (filter(EToH, id == ID & DATE <= date) %>%
     select(CODE))[1,2]
  )
#A Diagnoses of Rheumatoid Arthritis 
RA_loop <- as.character(
  (filter(RA, id == ID & DATE <= date) %>%
     select(CODE))[1,2]
  )
#According to FRAX this includes osteoporosis (M81, 7330) and diseases strongly associated with osteoporosis including type I (insulin dependent) diabetes (E10, 250), osteogenesis imperfecta in adults (Q78.0, 75651), untreated long-standing hyperthyroidism E05, 242)(these are the codes for hyperthyroidism irrespective of treatment), hypogonadism (E29.1, 2572 (male), E28.3, 25631, 25639 (female) including) premature menopause (<45 years), chronic malnutrition (intestinal malabsorption K90, 579, nutritional malabsorption E41-46, 261-263), or malabsorption and chronic liver disease
#An alternative to malabsorption could be to consider BMI <15
Osteoporosis_loop <- as.character(
  (filter(Osteoporosis, id == ID & DATE <= date) %>% 
     select(CODE))[1,2]
  )
#FRAX considers parental hip fracture a risk. We have substituted parental osteoporosis for this risk factor because there is a searchable diagnostic code
Parent_Osteoporosis_loop <- as.character(
  (filter(Parent_Osteoporosis, id == ID) %>% 
     select(CODE))[1,2]
  )
#Current exposure to oral glucocorticoids or exposure to oral glucocorticoids for more than 3 months at a dose eqivalent to prednisolone > 5mg daily
Glucocorticoid_data <- (
  filter(Glucocorticoid, id == ID & DATE <= date)
  )
Glucocorticoid_rf <- ifelse(Glucocorticoid_data[1,1] == 'NA', 'NA', Glucocorticoid_data[1,1])
#Exclude glucocorticoids taken for less than 3 months
#Add each risk factor to the 'CRF' tibble
CRF <- CRF %>% add_row(
ID = id,
Fracture = Fracture_loop,
Smoker = Smoker_loop,
EToH = EToH_loop,
RA = RA_loop,
Osteoporosis = Osteoporosis_loop,
Parent_Osteoporosis = Parent_Osteoporosis_loop,
Glucocorticoid_rf = as.character(Glucocorticoid_rf)
)
#Change the name of the tibble to calculate CRF score below so as to preserve the CRF tibble for retrospective analysis
CRF_calc <- CRF %>% filter(ID == id) %>% select(!ID)
#Convert risk factors to 0 (not present) or 1 (present)
CRF_calc[!is.na(CRF_calc)] <- '1'
CRF_calc[is.na(CRF_calc)] <- '0'
#Calculate CRF
CRF_score <- sum(as.double(CRF_calc))
#Put all the data you collected into 'EM_Dates'
EM_Dates = EM_Dates %>% add_row(
ID = id,
DATE = date,
CODE = CODE,
DESC = DESC,
HT.in = as.double(HT), 
WT.in = as.double(WT), 
Birth_Date = Birth_Date,
ageatEM = as.double(ageatEM),
ageatdeath = as.double(ageatdeath),
GENDER = as.character(Gender), 
Og_RACE = as.character(Race), 
CRF = as.double(CRF_score),
BMI = as.double(BMI)
)
}
#Save the original EM_Dates because the below chunks will make changes to it
EM_Dates_Og <- EM_Dates
#add time from fracture to death via mutate()
EM_Dates <- EM_Dates %>% mutate(time_to_death = ageatdeath - ageatEM)
#remove anyone < 50 years old at the first fracture. This should not include anyone in the cohort, but will catch anyone extra
EM_Dates <- EM_Dates %>% filter(ageatEM >= 50)
```
Calculate FRAX score below
How does this calculator work?
We have all the components needed to calculate FRAX score for each patient at the time of the sentinel event collected in EM_Dates. We will now create a tibble 'px' from EM_Dates in which BMI and age will be rounded and Race will be modified to match the descriptions in our FRAX paper charts. FRAX paper charts uses age, race, sex, BMI and # of CRFs to estimate FRAX risk. We will use 'join' to match each patient to a risk in the FRAX paper charts.
```{r FRAX calculator}
#add death date to EM_Dates and change the name of the tibble to 'px'
px <- Patients %>%
  select(ID, DEATH_DATE) 
px$ID <- as.double(px$ID) 
px <- px %>% right_join(EM_Dates, by = 'ID')
#Round BMI down to the nearest 5
floorfive <- function(x) {floor(x/5)*5}
px <- px %>% mutate(
  BMI = ifelse((BMI<45)&(BMI>15),floorfive(BMI), 
               ifelse (BMI>45,45,15)
               )
  )
#Round age down to the nearest 5
px <- px %>% mutate (
  AGE = ifelse((ageatEM<90)&(ageatEM>50),floorfive(ageatEM), 
               ifelse (ageatEM<50,50,90)
               )
  )
#Change race to white if not white, black, hispanic or asian
#Original race will be maintained the the column 'Og_RACE'
px$Og_RACE <- as.character(px$Og_RACE)
Races <- c("W","B","H","A")
px <- px %>% mutate(RACE = ifelse(
  (Og_RACE %in% Races), Og_RACE, "W")
  )

#Join FRAX and px to calculate 10 year risk of major osteoporotic fracture (do you mean major hip fracture?)
px <- px %>% 
  left_join(FRAX, by= c("GENDER","RACE","AGE","CRF","BMI"), copy=FALSE, keep=FALSE) %>%
  select(ID, Risk, Range_upper, Range_lower) %>% 
  arrange(ID) %>%
  na.omit()

EM_Dates_ed <- EM_Dates %>% left_join(px, by = 'ID')

write.csv(EM_Dates_ed, "C:\\Users\\acriscit\\Documents\\EM_Dates_ed.csv")
write.csv(CRF, "C:\\Users\\acriscit\\Documents\\CRF_EM_Dates_ed.csv")
```
#Descriptors
```{r descriptors}
percent_complete_cases <- (nrow(px)/nrow(EM_Dates))*100
percent_missing_FRAX <- (nrow(px)/nrow(EM_Dates_ed))*100
percent_missing_HT <- (sum(is.na(EM_Dates_ed$HT.in))/nrow(EM_Dates_ed))*100
percent_missing_WT <- (sum(is.na(EM_Dates_ed$WT.in))/nrow(EM_Dates_ed))*100

Flow <- tibble(percent_complete_cases = as.double(percent_complete_cases),
  percent_missing_FRAX = as.double(percent_missing_FRAX),
  percent_missing_WT = as.double(percent_missing_WT),
  percent_missing_HT = as.double(percent_missing_HT))

write.csv(Flow, "C:\\Users\\acriscit\\Documents\\Flow_ED.csv")

#Remove any tibbles you will not need
rm(EM_Dates, px, Flow)
#EM_Dates_ed contains all the info from the table
#EM_Dates_Og contains the original, unaltered info from the FRAX loop
```
#The below lines will select patients with a history of fracture. The patient pull in i2b2 stipulated that patients not have a Diagnoses of fracture prior to the index date (ED visit between 5/1/2013 and 5/31/2014). Importantly, patients may have a fracture after 5/31/2014. We want to capture these patients.

#It will also allow you to select patients without fracture from a population of patients with and without fractures.
```{r filtering}
#Select the ID of anyone with a non MOF Diagnoses: ICD9/10 codes- S62 hand, S92, S99 foot, M88 Osteodeformans/Paget's disease of bone, S02 Face/skull, V trauma due to a collision, Y04 assault by bodily force, C79.5 metastatic disease to bone
non_MOF <-  tibble_names$Diagnoses %>%
  #filter(grepl('[Ff]racture', DESC)) %>%
  filter(grepl('S62|S92|S99|M88|S02|V|Y04|C79.5', CODE)) %>%
  group_by(ID) %>%
  arrange(DATE, .by_group = TRUE) %>% 
  distinct(ID, .keep_all = TRUE) %>%
  select(ID, DATE, CODE, DESC)
### Not doing anything with the above info
```
#select patients who subsequently had a MOF
```{r MOF}
#Select anyone with MOF: hip fracture M84.359, M84.459, M84.659, Vertebral: 805|806|733.13|S12|M48.50|M80.08, Humerus (distal and shaft): 812.01|812.23|733.11|S42.2|M84.429A, Forearm: 813.45|813.8|733.12|813.0|S52|M84.429A
MOF <- tibble_names$Diagnoses %>%
  filter(grepl('[Ff]racture', DESC)) %>%
  filter(grepl('M84.359|M84.459|M84.659|820|S72.0|S72.90XA|S72.91XA|S72.92XA|805|806|733.13|S12|M48.50|M80.08|812.01|812.23|733.11|S42.2|M84.429A|813.45|813.8|733.12|813.0|S52|M84.429A', CODE)) %>%
  group_by(ID) %>%
  arrange(DATE, .by_group = TRUE) %>% 
  distinct(ID, .keep_all = TRUE) %>%
  select(ID, DATE, CODE, DESC)
MOF <- rename(MOF, fx_DATE = DATE, fx_CODE = CODE, fx_DESC = DESC)

#Join the date of fracture (after EM visit) to EM_Dates_ed
EM_Dates_ed$ID <- as.character(EM_Dates_ed$ID)
EM_Dates_ed <- EM_Dates_ed %>% left_join(MOF, by = 'ID')
#Create a column to see when the fracture occurred
EM_Dates_ed <- EM_Dates_ed %>% mutate(fx_time = difftime(fx_DATE, DATE,
         units = c("days")))
#Will be positive if after the ED visit
#Note that there should not be any fractures prior to 5/31/2014, if there are, i2b2 let patients slip through the collection criteria
num_pts_missed_i2b2 <- EM_Dates_ed %>% filter(fx_time < -1) %>% nrow()
#Overwrite prior ED_Dates_ed
write.csv(EM_Dates_ed, "C:\\Users\\acriscit\\Documents\\EM_Dates_ed.csv")
```
#Must discern which patients have diabetes within this cohort, and when they were diagnosed
```{r Diabetes calculator}
#Select patients with type 2 DM using ICD-10/9 code 'E11' or '250'
All_DM <- tibble_names$Diagnoses %>% 
  filter((grepl('E11|250', CODE))) %>% 
  #Select the ID and Date of first DM diagnosis in the record
  select(ID, DATE) %>% group_by(ID) %>%
  arrange(DATE, .by_group = TRUE) %>%
  distinct(ID, .keep_all = TRUE)
#Select patients with DM for whom we have a Frax score
ed <- EM_Dates_ed %>% filter(!is.na(Risk))
ID_DM <- inner_join(All_DM, ed, by = 'ID')
rm(All_DM, ed)

#calculate the time different between fracture and diagnosis of DM. A negative value indicates that the diagnosis was made after this visit
ID_DM <- ID_DM %>% mutate(
  days_before_ed = difftime(DATE.y, DATE.x, units = "days"))
ID_DM <- ID_DM %>% mutate(
  DM_Dx_postvisit = ifelse(days_before_ed < 0, "TRUE", "FALSE"))

#Pre-loop
Insulin <- tibble_names$Meds %>%
  filter(grepl('[Ii]nsulin', DESC)) %>%
  group_by(ID) %>%
  arrange(desc(DATE), .group_by = TRUE)
BG <- tibble_names$Labs %>%
  filter(grepl('GLU', CODE)) %>%
  filter(grepl('mg/dl|MG/DL', UNITS)) %>%
  filter(NUMERIC_VALUE != 0 & !is.na(NUMERIC_VALUE)) %>%
  arrange(desc(DATE))
HBA1C <- tibble_names$Labs %>%
  filter(grepl('HBA1C', CODE)) %>%
  filter(UNITS == '%') %>%
  arrange(desc(DATE))
```
Diabetes loop
Note that all characteristics are characterized prior to ED date, because we did the same for fracture date. For example, avg_HBA1C is the mean HBA1c in the year leading up to the ED visit.
```{r DM loop}
#Tibble to be filled by loop
DM_Dates_ed <- tibble (
  ID = numeric(),
  HBA1C_max = numeric(),
  HBA1C_max_date = character(),
  HBA1C_avg = numeric(),
  Insulin_start = character(),
  long_term_insulin = character(),
  BG = numeric(),
  BG_date = character(),
  BG1yavg = numeric(),
  BG1yr_num_values = numeric()
)

for(i in 1:nrow(ID_DM)) {
  id <- as.double(ID_DM$ID[i])
  
#Date of ED visit
  ed_date <- ID_DM$DATE.y[i]

  #date <- ID_DM$DATE.x[i]

#HBA1C's prior to fracture and date of measurement
HBA1C_loop <- HBA1C %>% 
  filter(ID == id) %>%
  filter(DATE <= ed_date)

HBA1C_max <- as.double(max(HBA1C_loop$NUMERIC_VALUE))
#HBA1C_max <- ifelse(is.na(HBA1C_max),0,HBA1C_max)

HBA1C_max_date <- (HBA1C_loop %>%
                     filter(NUMERIC_VALUE == HBA1C_max))$DATE[1]

#Average HBA1C in the year before fracture
HBA1C_avg <- HBA1C_loop$NUMERIC_VALUE %>% as.double() %>% mean()
#HBA1C_avg <- ifelse(is.na(HBA1C_avg),0,HBA1C_avg)

#Insulin and BG
Insulin_dates <- Insulin %>%
  filter(id == ID) %>%
  filter(DATE < ed_date)
Insulin_start <- Insulin_dates$DATE[1]
#Long term insulin use noted in problem list
long_term_insulin <- (Insulin_dates %>%
  filter(grepl('[Ll]ong term|[Ll]ong-term', DESC)))$DESC[1]
  
BGgen <- BG %>%
  filter(ID == id) %>%
  filter(DATE < ed_date)
BG_date <- BGgen$DATE[1]
BGnow <- as.double(BGgen$NUMERIC_VALUE)[1]
#BG averaged over 1 year prior to fracture if there are 2 or more values recorded
BG1yr <- BGgen %>% filter(DATE >= ed_date - years(1))
BG1yr_num_values <- BG1yr %>% nrow()
BG1yavg <- ifelse(BG1yr_num_values > 1, BG1yr$NUMERIC_VALUE %>% as.double() %>% mean(), NA)

rm(BG1yr)

  DM_Dates_ed <- DM_Dates_ed %>% add_row(
   ID = id,
   HBA1C_max = as.double(HBA1C_max),
   HBA1C_max_date = as.character(HBA1C_max_date),
   HBA1C_avg = as.double(HBA1C_avg),
   Insulin_start = as.character(Insulin_start),
   long_term_insulin = long_term_insulin,
   BG = BGnow,
   BG_date = as.character(BG_date),
   BG1yavg = BG1yavg,
   BG1yr_num_values = BG1yr_num_values
)
  rm(HBA1C_loop)
  rm(Insulin_dates)
  rm(BGgen)
  rm(BG1yr)
}

#replace 'NaN' with NA
DM_Dates_ed$HBA1C_avg[DM_Dates_ed$HBA1C_avg == "NaN"] <- NA
#Make long term insulin true or false
DM_Dates_ed$long_term_insulin <- ifelse(!is.na(DM_Dates_ed$long_term_insulin), "TRUE", "FALSE")
#Join the info pulled into 'ID_DM' form FRAX loop
ID_DM$ID <- as.double(ID_DM$ID)
DM_Dates_ed <- DM_Dates_ed %>%
  left_join(ID_DM, by='ID')
#Rename the date columns
DM_Dates_ed <- DM_Dates_ed %>% rename(ED_DATE = DATE.y, DM_DATE = DATE.x)

write.csv(DM_Dates_ed, "C:\\Users\\acriscit\\Documents\\DM_Dates_ed.csv")

rm(ID_DM)
rm(BG)
rm(Insulin)
rm(HBA1C)
```
Visits
run the loop below
```{r visit_loop}

#Information about number of visits and date of visits relative to fracture date
Visit_Dates <-  tibble(
  ID = numeric(),
  first_visit_date = mdy(),
  last_visit_date = mdy(),
  num_visits_before_ed = numeric(),
  num_visits_after_ed = numeric(),
  CCI = numeric ()
)

for (i in 1:nrow(EM_Dates_ed)){
  id <- as.double(EM_Dates_ed$ID[i])

date <- (EM_Dates_ed %>% 
    filter(ID == id) %>%
    select(ID, DATE))[[1,2]]

#Create a tibble with one patients into
visit <- Visits %>% 
    filter(ID == id)

first_visit_date <- visit$START_DATE[1]
last_visit_date <- (
  visit %>% 
    arrange(desc(START_DATE))
  )$START_DATE[1]

num_visits_before_ed <- visit %>% 
  filter(START_DATE < date) %>%
  #mutate(START_DATE = floor_date(START_DATE, "day")) %>%
  #group_by(START_DATE) %>% 
  #distinct(START_DATE, .keep_all = TRUE) %>%
  nrow()
num_visits_after_ed <- visit %>% 
  filter(START_DATE > date) %>% 
  #mutate(START_DATE = floor_date(START_DATE, "day")) %>%
  #group_by(START_DATE) %>% 
  #distinct(START_DATE, .keep_all = TRUE) %>%
  nrow()
#LOS_fx <- visit %>% 
  #filter(TYPE == 'I|S') %>% 
 # filter(START_DATE <= date & END_DATE >= date) %>%
 # select(LOS) %>% as.double()[[1]]

#Finds the most recent CCI prior to the fracture date
CCI <- visit %>%
  filter(!is.na(CCI) & START_DATE < date) %>%
 # filter(START_DATE <= date) %>%
  arrange(desc(START_DATE)) %>%
  distinct(ID, .keep_all = TRUE) %>%
  select(CCI)
CCI <- as.double(CCI[1,2])

Visit_Dates <- Visit_Dates %>% add_row(
  ID = id,
  first_visit_date = first_visit_date,
  last_visit_date = last_visit_date,
  num_visits_before_ed = num_visits_before_ed,
  num_visits_after_ed = num_visits_after_ed,
  CCI = as.double(CCI)
)
rm(visit)
}

write.csv(Visit_Dates, "C:\\Users\\acriscit\\Documents\\Visit_Dates_ed.csv")
rm(Visit_Dates)
#Join Visits to tibble_names and save tibble_names as a patient database for future use
#remove 'test' once you've tested this
test <- append(tibble_names, list(Visits))
rm(Visits)

```
Osteo meds
```{r osteo meds}
#FDA approved osteoporosis medication usage
osteo_med_all <- tibble_names$Meds %>% 
  filter(
    grepl('[Ee]tidronate|[Aa]lendronate|[Ii]bandronate|[Zz]oledronic|[Rr]isedronate|[Cc]alcitonin|[Dd]enosumab|[Ee]strogen|[Rr]aloxifene|[Bb]azodoxifene|[Rr]omosozumab|[Tt]eriparatide|[Aa]baloparatide', DESC) &
    !grepl('[Rr]eceptor', DESC)
         ) %>%
  filter(MODIFIER == "mg") %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)
```
Osteo loop
```{r osteo_loop}
Osteo_Meds <- tibble(
  ID = numeric(),
  osteo_med_start_date = character(),
  osteo_med_start_dose_mg = numeric(),
  osteo_med_last_date = character(),
  osteo_med_name = character()
)

for (i in 1:nrow(EM_Dates_ed)){
  id <- as.double(EM_Dates_ed$ID[i])

date <- (EM_Dates_ed %>% 
    filter(ID == id) %>%
    select(ID, DATE))[[1,2]]
  
  #FDA approved osteoporosis medication usage
osteo_med <- (filter(osteo_med_all, id == ID))
osteo_med_name <- as.character(osteo_med$DESC[1])
osteo_med_start_date <- osteo_med$DATE[1]
osteo_med_start_dose <- osteo_med$UNITS[1]
osteo_med_last_date <- (osteo_med %>%
  arrange(desc(DATE)))$DATE[1]

Osteo_Meds <- Osteo_Meds %>% add_row(
  ID = id,
  osteo_med_start_date = as.character(osteo_med_start_date),
  osteo_med_start_dose_mg = as.double(osteo_med_start_dose),
  osteo_med_last_date = as.character(osteo_med_last_date),
  osteo_med_name = as.character(osteo_med_name)
)
rm(osteo_med)
}

Osteo_Meds <- na.omit(Osteo_Meds)
write.csv(Osteo_Meds, "C:\\Users\\acriscit\\Documents\\Osteo_Meds_ed.csv")
rm(Osteo_Meds)