---
title: "FRAX calculator for diabetic patients without fracture"
author: "Anthony Criscitiello and Ellen Quillen"
date: "1/10/2021"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "FRAX calculator for I2b2 results"
output: html_document
---
```{r setup}
library(tidyverse)
library(readxl)
library(readr)
library(lubridate)
library(tibble)
library(purrr)
```
^I find it helpful to set everything up at the beggining and you can also make some modifications to how this Rmd turns into an html. Here is cheatsheet on options for R markdown: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
##Introduction
```{r load FRAX ppr charts and patient data}
#load FRAX paper charts (10-year risk of major osteoporotic fracture based on BMI)
FRAX <-  read_csv("C:\\Users\\acriscit\\Documents\\FRAX.csv", col_types = "ccdddddd")
names(FRAX)<- c("GENDER", "RACE", "AGE","CRF","BMI","Risk", "Range_upper", "Range_lower")

Data_Request_1501_dem_01_06_2021 <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_dem_01_06_2021.csv")
Data_Request_1501_dx_01_06_2021 <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_dx_01_06_2021.csv")
Data_Request_1501_fam_hx_01_06_2021 <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_fam_hx_01_06_2021.csv")
Data_Request_1501_img_01_06_2021 <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_img_01_06_2021.csv")
Data_Request_1501_Labs_01_07_2021 <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_Labs_01_07_2021.csv")
Data_Request_1501_meds_01_06_2021 <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_meds_01_06_2021.csv")
Data_Request_1501_vitals_01_11_2021 <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_vitals_01_11_2021.csv")

#Label each list element with it's associated tab from the excel sheet
tab_names <- excel_sheets(path = FRAX_data)
#Below is the slow line
FRAX_data <- map(tab_names, ~read_excel(path = FRAX_data, sheet = .x))
FRAX_data <- FRAX_data %>% set_names(tab_names)
```
Now that the data is loaded into R we will start to process it into a more readable ('tidy') form
```{r tidy_frax}

#Collect the tabs from 'xl_data' that contain info used to calculate FRAX score
tab_names2 <- tibble(
Row = (1:length(tab_names)),
tab_names = tab_names
)

Frax_tabs <- tab_names2 %>% 
    filter(
    grepl('Diagnoses', tab_names) | 
    grepl('Labs', tab_names) | 
    grepl('Meds', tab_names) | 
    grepl('Vitals', tab_names) | 
    grepl('Smoking', tab_names)
  )
FRAX_pts <- FRAX_data[c(Frax_tabs$Row)]

#Selecting specific colums below. This will make tabs uniform and mergeable
for (i in 1:length(FRAX_pts)){
 FRAX_pts[[i]] <- FRAX_pts[[i]] %>% select(
  contains(
    c(
      'PATIENT',
      'VISIT',
      'DATE',
      'CODE',
      'DESC', 
      'TEXT', 
      'NUMERIC', 
      'UNITS',
      'MODIFIER'
      )
    )
  )
}

#rename columns and remove spaces from column names
col_names <- c(
  "ID", 
  "VISIT", 
  "DATE",
  "CODE",
  "DESC", 
  "TEXT_VALUE", 
  "NUMERIC_VALUE", 
  "UNITS",
  "MODIFIER"
)
for (i in 1:length(FRAX_pts)){
  names(FRAX_pts[[i]]) <- col_names[1:ncol(FRAX_pts[[i]])]
}

#Tidy the data by condensing the 'list' into a 'tibble' using 'bind_rows'
#BEWARE USING X AND Y OR SIMILARLY NON-DESCRIPT VARIABLES FOR OBJECTS YOU NEED TO USE OVER A LONG PERIOD. MORE DESCRIPTIVE NAMES HELP YOU REMEMBER WHAT YOU WERE DOING DOWN THE LINE
x <- bind_rows(FRAX_pts)
x$DATE <- mdy_hm(x$DATE, quiet = TRUE
)

#Collect the tabs containing demographic info used to calculate FRAX score from 'FRAX_data'
info_tabs <- tab_names2 %>% filter(grepl('Patients', tab_names))
FRAX_info <- FRAX_data[c(info_tabs$Row)]

#Rename columns
#coly <- colnames(FRAX_data$Patients)
col_names <- c(
  "ID", 
  "AGE_NOW", 
  "AGE_AT_DEATH",
  "BIRTH_DATE",
  "DEATH_DATE",
  "MATCH",
  "GENDER", 
  "RACE", 
  "MIXED RACE", 
  "ETHNICITY"
)

for (i in 1:length(FRAX_info)){
  names(FRAX_info[[i]]) <- col_names[1:ncol(FRAX_info[[i]])]
}

y <- bind_rows(FRAX_info)

y$BIRTH_DATE <- mdy_hm(y$BIRTH_DATE, quiet = TRUE
)
```
Now that the patient data is more readable, we will create a tibble that contains data needed to calculate a FRAX score for each patient. This will primarily be accomplished with the use of a loop.
```{r calc_frax}

#Create a tibble ('Current_Dates') with columns for each component of the FRAX score
Current_Dates <- tibble(
  ID = numeric(),
  DATE = mdy_hms(), #last visit
  HT.in = numeric(), 
  WT.in = numeric(),
  AGE = numeric(), 
  GENDER = character(), 
  Og_RACE = character(), 
  CRF = numeric(), #CRF is Combined Risk Factors
  BMI = numeric()
)

#Create a tibble with all the IDs of patients who have had a fracture
no_fracs <- x %>%
  filter(!grepl('[Ff]racture', DESC)) %>% 
  select(ID, DATE, TEXT_VALUE) %>% group_by(ID) %>% arrange(DATE, .by_group = TRUE)
IDs <- no_fracs %>% distinct(ID, .keep_all = TRUE)
```
The below loop is will collect and distill information relevant to calculating FRAX score. Each cycle of this loop will collect data for 1 patient
```{r calc_frax_loop}

for (i in 1:nrow(IDs)){
  id <- as.double(IDs$ID[i])

date <- (IDs %>% 
    filter(ID == id) %>% arrange(desc(DATE)))$DATE[1]

#Most recent height and weight
HT <- (x %>% 
         filter(CODE == "HT", ID == id) %>% 
         arrange(desc(DATE)) %>% 
         select(NUMERIC_VALUE)
       )[1,1]
WT <- (x %>% 
         filter(CODE == "WT", ID == id) %>% 
         arrange(desc(DATE)) %>% 
         select(NUMERIC_VALUE)
       )[1,1]

#Calculate BMI
BMI <- (703*as.double(WT)/as.double(HT)^2)

#Calcuate age at time of fracture
yID <- y %>% filter(ID == id)

age <- yID$AGE_NOW[1]

Gender <- yID %>% select(GENDER)

Race <- yID %>% select(RACE)

#Tibble where the risk factors will be compiled
CRF <- tibble(
  Fracture = character(),
  Smoker = character(),
  EToH = character(),
  RA = character(),
  Osteoporosis = character(),
  Parent_Osteoporosis = character(),
  Glucocorticoid = character()
)
#Collect risk factors into above tibble from  (Fracture  Parent fracture Smoking Glucocorticoids RA  Alcohol use Secodary osteoporosis)
#If there is no observation prior to DATE, it is assumed that the risk factor was not present. For example, if there is no prior fracture, it is assumed that there was no prior fracture
ID_data <- x %>% filter(ID == id & DATE < date)
#'[Ff]racture'
Fracture <- as.character(
  (ID_data %>% 
     filter(grepl('Fracture|fracture', DESC)) %>% select(DESC))[1,1])
Smoker <- as.character(
  (ID_data %>% 
     filter(grepl('Smoker|smoker', DESC)) %>% select(DESC))[1,1])
EToH <- as.character(
  (ID_data %>% 
     filter(grepl('Alcoholism|alcoholism', DESC)) %>% select(DESC))[1,1])
RA <- as.character(
  (ID_data %>% 
     filter(grepl('Rheumatoid|rheumatoid', DESC),grepl('M05|M06|714', CODE)) %>% select(DESC))[1,1])
Osteoporosis <- as.character(
  (ID_data %>% 
     filter(grepl('Osteoporosis|osteoporosis', DESC), grepl('M80|733', CODE)) %>% select(DESC))[1,1])
#This should be parent hip fracture
Parent_Osteoporosis <- as.character(
  (ID_data %>% 
     filter(grepl('Z82.62|V17.81', CODE)) %>% select(CODE))[1,1])
#Current exposure to oral glucocorticoids or exposure to oral glucocorticoids for more than 3 months at a dose eqivalent to prednisolone > 5mg daily
Glucocorticoid <- as.character(
  Glucocorticoid <- (
  (x %>% 
     filter(DATE > (date - months(3)) & UNITS == 'mg' & grepl('sone| solone', DESC) & grepl('[Ii]njection| [Tt]ablet| [Ii]vbp', DESC)
     )))[1,1])

#Put this table into use to ensure dose is adequate
#steriod_conversion_table <- tibble(
#  steriod = c("cortisone", "hydrocortisone", "Methylprednisone", "Prednisolone", "Prednisone", "Triamcinolone", "Betamethasone", "Dexamethasone"),
#  Dose = c(5, 4, 0.8, 1, 1, 0.8, 0.12, 0.15)
#)
#To parse this out use ifelse statements. If it is injection/iv count it (ifelse(filter(grepl('IV or injection'))), 1, 0
#separate this column into multiple columns using tidy data to pivot
#These lines are held because they slow down the loop

#Add each risk factor to the 'CRF' tibble
#Note that there is no ID in this tibble. This is because we have selected data specific to one patient in line 188. Furthermore, we are still within the loop. Each cycle of a loop collects data for one patient
CRF <- CRF %>% add_row(
Fracture = Fracture,
Smoker = Smoker,
EToH = EToH,
RA = RA,
Osteoporosis = Osteoporosis,
Parent_Osteoporosis = Parent_Osteoporosis,
Glucocorticoid = Glucocorticoid
)
#Convert risk factors to 0 (not present) or 1 (present)
CRF[!is.na(CRF)] <- '1'
CRF[is.na(CRF)] <- '0'
#Calculate CRF
CRF_score <- sum(as.double(CRF))

#Put all the data you collected into 'Fracture_Dates'
Current_Dates = Current_Dates %>% add_row(
ID = id,
DATE = date,
HT.in = as.double(HT), 
WT.in = as.double(WT), 
AGE = as.double(age), 
GENDER = as.character(Gender), 
Og_RACE = as.character(Race), 
CRF = as.double(CRF_score),
BMI = as.double(BMI)
)
}
#end of loop
```
The tibble 'Fracture_Dates' now contains a row for each patient and associated information necessary to calculate FRAX score. Below we will use the 'FRAX' tibble create from publicly available FRAX paper charts (https://www.sheffield.ac.uk/FRAX/charts.aspx) to calculate the FRAX score for each patient.

Discern whether the patients with fracture have diabetes.
What is the duration of disease prior to fracture?
Were they well controlled?
```{r Diabetes calculator}
DM_dates <- tibble (
  ID = numeric(),
  Dx_DATE = mdy_hms(),
  Duration = duration(),
  HBA1C_max = numeric(),
  HBA1C_max_date = character(),
  HBA1C_avg = numeric()
)

#Select patients with DM
All_DM <- x %>% 
  filter((grepl('[Dd]iabetes', DESC))) %>% 
  select(ID, DATE) %>% group_by(ID) %>% arrange(DATE, .by_group = TRUE)

ID_DM <- All_DM %>% distinct(ID, .keep_all = TRUE) %>% distinct(ID, .keep_all = TRUE)
```

```{r loopDM}
for(i in 1:nrow(ID_DM)) {
  id <- as.double(ID_DM$ID[i])
  
  DM_Dx_date <- (ID_DM %>% 
     filter(ID == id))$DATE[1]
  
  date <- (Current_Dates %>% filter(ID == id))$DATE[1]
  DM_duration <- date - DM_Dx_date
  

#HBA1C's prior to fracture and date of measurement
HBA1C <- x %>% 
  filter(ID == id) %>%
  filter(grepl('HBA1C', CODE)) %>%
  filter(UNITS == '%')

HBA1C_max <- as.double(max(HBA1C$NUMERIC_VALUE))
HBA1C_max <- ifelse(is.na(HBA1C_max),0,HBA1C_max)

HBA1C_max_date <- (HBA1C %>%
                     filter(NUMERIC_VALUE == HBA1C_max))$DATE[1]

#Average HBA1C over the last year
HBA1C_avg <- HBA1C$NUMERIC_VALUE %>% as.double() %>% mean()
HBA1C_avg <- ifelse(is.na(HBA1C_avg),0,HBA1C_avg)

  DM_dates <- DM_dates %>% add_row(
   ID = id,
   Dx_DATE = DM_Dx_date,
   Duration = DM_duration,
   HBA1C_max = as.double(HBA1C_max),
   HBA1C_max_date = as.character(HBA1C_max_date),
   HBA1C_avg = as.double(HBA1C_avg)
)
}

write.csv(DM_dates, "C:\\Users\\acriscit\\Documents\\DM_dates.csv")

```
#Calculate FRAX score below
```{r calc_frax}
px <- Current_Dates

#Round BMI down to the nearest 5
floorfive <- function(x) {floor(x/5)*5}
px <- px %>% mutate(
  BMI = ifelse((BMI<45)&(BMI>15),floorfive(BMI), 
               ifelse (BMI>45,45,15)
               )
  )

#Round age down to the nearest 5
px <- px %>% mutate (
  AGE = ifelse((AGE<90)&(AGE>50),floorfive(AGE), 
               ifelse (AGE<50,50,90)
               )
  )

#Change race to white if not white, black, hispanic or asian
#Original race will be maintained the the column 'Og_RACE'
px$Og_RACE <- as.character(px$Og_RACE)
Races <- c("W","B","H","A")
px <- px %>% mutate(RACE = ifelse(
  (Og_RACE %in% Races), Og_RACE, "W")
  )

#Join FRAX and px to calculate 10 year risk of major osteoporotic fracture
fx <- px %>% left_join(FRAX, by= c("GENDER","RACE","AGE","CRF","BMI"), copy = FALSE, keep = FALSE)%>%
select(ID, Risk, Range_upper, Range_lower) %>% 
  arrange(ID)

percent_complete_cases <- (sum(complete.cases(fx))/nrow(fx))*100

#Remove incomplete cases
fxfinal <- na.omit(fx)

write.csv(fxfinal, "C:\\Users\\acriscit\\Documents\\fxfinal.csv")
write.csv(Current_Dates, "C:\\Users\\acriscit\\Documents\\Current_Dates.csv")

```
Other questions to ask by modifying this script
  Current question: Does FRAX estimate fracture risk for DM and non-DM patients equally? and Do FRAX scores differ between those with and without diabetes?
Others:
  Does duration of diabetes alter fracture risk?
  Does FRAX score change leading up to a fracture?



