---
title: "FRAX calculator for I2b2 results 1/10/21"
author: "Anthony Criscitiello and Ellen Quillen"
date: "1/10/2021"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "FRAX calculator for I2b2 results"
output: html_document
---
```{r setup}
library(tidyverse)
library(readxl)
library(lubridate)
library(tibble)
library(purrr)
```
^I find it helpful to set everything up at the beggining and you can also make some modifications to how this Rmd turns into an html. Here is cheatsheet on options for R markdown: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
##Introduction
The goal of this is to .. (anything in white will show up as standard text, anything in gray will be analyzed as script)
You can decide how many chunks you want to break your script into. For me this is generally determined by the amount of related information I want to include.
I put my comments below in upper case to try and make them easier for you to find.
```{r calc_frax}
#load FRAX paper charts (10-year risk of major osteoporotic fracture based on BMI)
#FRAX <-  read_csv("E:/Dropbox/Students/Anthony/FRAX.csv", header = TRUE, stringsAsFactors = FALSE)
FRAX <-  read_csv("C:\\Users\\acriscit\\Documents\\FRAX.csv", col_types = "ccdddddd")
names(FRAX)<- c("GENDER", "RACE", "AGE","CRF","BMI","Risk", "Range_upper", "Range_lower")
#Load I2b2 data
#xl_data <- "E:/Dropbox/Students/Anthony/20201020180719000_acriscit.xlsx"
xl_data <- "C:\\Users\\acriscit\\Documents\\20201020180719000_acriscit.xlsx"
#Label data by excel tab
tab_names <- excel_sheets(path = xl_data)
#The below line is very slow
# CONSIDER USING THIS SUBSTITUT IN purrr (PART OF THE TIDYVERSE):https://jtr13.github.io/spring19/ss5593&fq2150.html
pts <- map(tab_names, ~read_excel(path = xl_data, sheet = .x))
#only slightly faster than lapply if at all. The lag is probably caused by the read_excel function
pts <- pts %>% set_names(tab_names)
```
#Combine tabs that contain the same type of information
  #Look at the excel sheet to see which data is stored across     multiple tabs and double check that the code covers these
  #Create a tibble of tab names and numbers
tab_names2 <- tibble(
Row = (1:length(tab_names)),
tab_names = tab_names
)
  #Meds
#Select all tabs that contain meds
Med_tabs <- tab_names2 %>% filter((grepl('Meds', tab_names)))
#ASSUMING THAT THESE TABS ARE ALWAYS IDENTICAL, THIS IS FINE
#HOWEVER, I WOULD SUGGEST THAT YOU INCLUDE AN IF STATEMENT TO CHECK THAT THE NAMES ARE THE SAME BETWEEN THE TABS
#ARE THESE ACUALLY BIND?
#Combine these tabs into one
n <- ncol(pts$Meds)
pts$Meds <- bind_rows(pts[Med_tabs$Row])
#ifelse statement to ensure that the bound tabs were identical
ifelse(ncol(pts$Meds)==n, , "STOP: the Med_tabs were not identical")
#Repeat with all other tabs that contain the same type of info
  #Labs
Lab_tabs <- tab_names2 %>% filter((grepl('Labs', tab_names)))
n <- ncol(pts$Labs)
pts$Labs <- bind_rows(pts[Lab_tabs$Row])
ifelse(ncol(pts$Labs)==n, , "STOP: the Lab_tabs were not identical")
  #Vitals
Vitals_tabs <- tab_names2 %>% filter((grepl('Vital', tab_names)))
n <- ncol(pts$Vitals)
pts$Vitals <- bind_rows(pts[Vitals_tabs$Row])
ifelse(ncol(pts$Vitals)==n, , "STOP: the Vitals_tabs were not identical")
#Remove the tabs that were merged
tabs_to_remove <- tab_names2 %>% filter((grepl('_p', tab_names)))
#Ensure that the right number of tabs were joined and removed
tabs <- list(Vital_tabs, Med_tabs, Lab_tabs)
ifelse(nrow(tabs_to_remove) == (sapply(tabs, function(x) nrow(x)) %>% sum())-3,,"STOP:the incorrect number of tabs will be removed")
n <- length(pts)
pts <- pts[-c(tabs_to_remove$Row)]
ifelse(nrow(tabs_to_remove)+length(pts) == n,,"STOP: An incorrect number of tabs were removed")
#AGAIN, THIS IS FINE, ASSUMING THAT THE COLUMN NAMES ARE ALWAYS GOING TO BE THE SAME
#THIS WILL NEED TO BE CHECKED MANUALLY FOR ANY FUTURE DATA
```{r calc_Frax_cont}
#Select the tabs used to calculate FRAX score
tab_names2 <- tibble(
Row = (1:length(tab_names)),
tab_names = tab_names
)
Frax_tabs <- tab_names2 %>% 
  filter(
  grepl('Diagnoses', tab_names) | 
  grepl('Labs', tab_names) | 
  grepl('Meds', tab_names) | 
  grepl('Vitals', tab_names) | 
  grepl('Smoking', tab_names)
  )
FRAX_data <- pts[c(Frax_tabs$Row)]
#rename columns and to remove spaces from column names
col_names <- c("ID", "VISIT", "DATE","CODE", "DESC", "MODIFIER", "TEXT_VALUE", "NUMERIC_VALUE", "UNITS")
for (i in 1:length(FRAX_data)){
  names(FRAX_data[[i]]) <- col_names[1:ncol(FRAX_data[[i]])]
}
#Tidy the data by binding rows
#IS THIS JUST A SUBSET/FILTER CALL? IT DOESN'T SEEM LIKE YOU ARE BINDING ROWS
#BEWARE USING X AND Y OR SIMILARLY NON-DESCRIPT VARIABLES FOR OBJECTS YOU NEED TO USE OVER A LONG PERIOD. MORE DESCRIPTIVE NAMES HELP YOU REMEMBER WHAT YOU WERE DOING DOWN THE LINE
x <- bind_rows(FRAX_data)
x$DATE <- mdy_hms(x$DATE)

info_tabs <- tab_names2 %>% filter(grepl('Patients', tab_names))
FRAX_info <-pts[c(info_tabs$Row)]
names(FRAX_info)<-c("ID", "AGE_NOW", "AGE_AT_DEATH", "GENDER", "RACE", "MIXED RACE", "ETHNICITY")
y <- bind_rows(FRAX_info)

#Nothing from the visit tabs will be used for the FRAX calculator so it can be ignored along with any other data that does not pass the filter
visit_tabs <- tab_names2 %>% filter(grepl('Visits', tab_names))
FRAX_visits <- pts[c(visit_tabs)]
names(FRAX_visits)<- c("ID", "VISIT", "DATE", "END_DATE", "CODE")

#Create a tibble ('Fracture_Dates') with the date of the most recent fracture for each patient and associated FRAX characteristics
Fracture_Dates <- tibble(
ID = numeric(),
DATE = mdy_hms(),
CODE = character(),
DESC = character(),
HT.in = numeric(), 
HTDaysBeforeFrac = numeric(), 
WT.in = numeric(), 
WTDaysBeforeFrac = numeric(), 
AGE = numeric(), 
GENDER = character(), 
RACE = character(), 
CRF = numeric(),
#CRF is Combined Risk Factors
BMI = numeric(),
Risk = numeric(),
Range_upper = numeric(),
Range_lower = numeric()
)
#Isolate patients with pathologic fracture by ID
all_fracs <- x %>% filter((grepl('M80', CODE))) %>% filter((grepl('fracture', DESC))) %>% select(ID)
all_IDs <- tibble(y['ID'])
ID_fracs <- right_join(all_IDs, all_fracs, by = NULL) %>% distinct()
#loop that will create a tibble with info about patient fractures
  #Use patient IDs from ID_fracs to run the loop
for (i in 1:nrow(ID_fracs)){
  id <- as.double(ID_fracs[i,])
#Filtering by ICD-10 code 'M80' (pathologic fracture) and diagnosis described as 'fracture' to select the first fracture a patient had
#For patients with multiple fractures you'll need to add another line that selects 2nd fractures
#I'VE REFORMATTED THIS. IT IS TOTALLY OPTIONAL BUT I FIND IT MAKES IT EASIER TO READ
pt_fractures <- (x %>% 
        filter(ID == id) %>% 
        filter((grepl('M80', CODE))) %>% 
        filter((grepl('fracture', DESC))) %>% 
        arrange(DATE))[1,]
date <- (pt_fractures %>% select(DATE))[1,1]
#Height and weight closest to the time of fracture
#Add an ifelse statement to collect HT and WT for pts without measurements prior to fracture
#HTWT <- (x %>% filter(CODE == "HT"|CODE == "WT", ID == id) %>% ifelse(filter(DATE <= date), ) %>% arrange(desc(DATE))
HT <- (x %>% filter(CODE == "HT", ID == id, DATE <= date) %>% arrange(desc(DATE)) %>% select(NUMERIC_VALUE))[1,1]
WT <- (x %>% filter(CODE == "WT", ID == id, DATE <= date) %>% arrange(desc(DATE)) %>% select(NUMERIC_VALUE))[1,1]
#Calculate BMI
BMI <- (703*as.double(WT)/as.double(HT)^2)
#Calcuate age at time of fracture
ageatfracture <- as.double(y %>% filter(ID == id) %>% select(AGE_NOW)) - (as.double(year(today())) - as.double(year(Fracture_Dates$DATE[i])))
#Select gender
Gender <- y %>% filter(ID == id) %>% select(GENDER)
#Select race
Race <- y %>% filter(ID == id) %>% select(RACE)
#Tibble where CRFs will be compiled
CRF <- tibble(
  Fracture = character(),
  Smoker = character(),
  EToH = character(),
  RA = character(),
  Osteoporosis = character(),
  Parent_Osteoporosis = character(),
  Glucocorticoid = character()
)
#Select CRFs (Fracture  Parent fracture Smoking Glucocorticoids RA  Alcohol use Secodary osteoporosis)
#ARE YOU TRACKING THE DIFFERENCE BETWEEN DATE AND date?
#RELATED, WHAT HAPPENS IF THERE ISN'T A PRIOR, RECENT OBSERVATION
e <- x %>% filter(ID == id, DATE < date)
#THIS REGEX SHOUP ALSO WORK: '[Ff]racture'
Fracture <- as.character((e %>% filter(grepl('Fracture|fracture', DESC)) %>% select(DESC))[1,1])
Smoker <- as.character((e %>% filter(grepl('Smoker|smoker', DESC)) %>% select(DESC))[1,1])
EToH <- as.character((e %>% filter(grepl('Alcoholism|alcoholism', DESC)) %>% select(DESC))[1,1])
RA <- as.character((e %>% filter(grepl('Rheumatoid|rheumatoid', DESC),grepl('M05|M06|714', CODE)) %>% select(DESC))[1,1])
Osteoporosis <- as.character((e %>% filter(grepl('Osteoporosis|osteoporosis', DESC), grepl('M80|733', CODE)) %>% select(DESC))[1,1])
Parent_Osteoporosis <- as.character((e %>% filter(grepl('Z82.62|V17.81', CODE)) %>% select(CODE))[1,1])
Glucocorticoid <- as.character((e %>% filter(DATE < date, grepl('sone|solone', DESC)) %>% select(DESC))[1,1])
#Work on glucs filter criteria and find out whether they need to be taking it at the time of fracture for it to count
#THIS TIBBLE DOESN'T HAVE A PATIENT ID IN IT? ARE YOU MAKING A NEW TIBBLE FOR EACH PT? 
CRF <- CRF %>% add_row(
Fracture = Fracture,
Smoker = Smoker,
EToH = EToH,
RA = RA,
Osteoporosis = Osteoporosis,
Parent_Osteoporosis = Parent_Osteoporosis,
Glucocorticoid = Glucocorticoid
)
#Convert risk factors to 0 (not present) or 1 (present)
CRF[!is.na(CRF)] <- '1'
CRF[is.na(CRF)] <- '0'
#Calculate CRF
CRF_score <- sum(as.double(CRF))
#Put all the data you collected into 'Fracture_Dates'
Fracture_Dates = Fracture_Dates %>% add_row(
ID = id,
DATE = z$DATE,
CODE = z$CODE,
DESC = z$DESC,
HT.in = as.double(HT), 
WT.in = as.double(WT), 
AGE = as.double(ageatfracture), 
GENDER = as.character(Gender), 
RACE = as.character(Race), 
CRF = as.double(CRF_score),
BMI = as.double(BMI)
)
}
#end of loop
#For some reason using filter with dates creates an error, however, these functions appear to work as expected
#px will sustain alterations in order to calculate FRAX score
px <- Fracture_Dates
#Do they have DMII at the time of fracture? How long before?
#Calculate FRAX score below
#Round BMI down to the nearest 5
floorfive <- function(x) {floor(x/5)*5}
px <- px %>% mutate(BMI = ifelse((BMI<45)&(BMI>15),floorfive(BMI), ifelse (BMI>45,45,15)))
#Round age down to the nearest 5
px <- px %>% mutate (AGE = ifelse((AGE<90)&(AGE>50),floorfive(AGE), ifelse (AGE<50,50,90)))
#I'VE CHANGED NOMENCLATURE TO WHITE AS THAT IS WHAT IS IN THE DATASET
#Change race to white if not white, black, hispanic or asian
#THIS MAKES SENSE TO DO, BUT WE NEED A WAY TO FLAG ANYONE WHO IS CHANGE THIS WAY, PERHAPS WITH A NEW  COLUMN TO REPRESENT ORIGINAL RACE
px$RACE <- as.character(px$RACE)
Races <- c("W","B","H","A")
px <- px %>% mutate(RACE = ifelse((RACE %in% Races), RACE, "W"))
#Join FRAX and px to calculate 10 year risk of major osteoporotic fracture
fx <- px %>% left_join(FRAX, by= c("GENDER","RACE","AGE","CRF","BMI"), copy = FALSE, keep = FALSE) %>% select(ID, Risk, Range_upper, Range_lower) %>% arrange(ID)
#Add FRAX score to 'Fracture_Dates'
Fracture_Dates = Fracture_Dates %>% add_row(
Risk = fx$Risk,
Range_upper = fx$Range_upper,
Range_lower = fx$Range_lower)
#or
write.csv(fx, "E:/Dropbox/Students/Anthony/fx.csv")
```
A final thought that occurred to me is that we may want to calculate a FRAX score at every time we see them rather than just the time prior (not necessarily in this script, just in general). We know that duration of diabetes is a risk factor for fracture in diabetics, it would be interesting to know how duration of diabetes is associated with fracture risk. 



