---
title: "FRAX calculator for Diabetic patients with fracture"
author: "Anthony Criscitiello and Ellen Quillen"
date: "1/10/2021"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
```

---
title: "FRAX calculator for I2b2 results"
output: html_document
---
```{r setup}
library(tidyverse)
library(readxl)
library(readr)
library(lubridate)
library(tibble)
library(purrr)
library(stringr)

#I find it helpful to set everything up at the beggining and you can also make some modifications to how this Rmd turns into an html. Here is cheatsheet on options for R markdown: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
##Introduction
#load FRAX paper charts (10-year risk of major osteoporotic fracture based on BMI)
FRAX <-  read_csv("H:/Hip fracture/FRAX.csv", col_types = "dccdddddd")
FRAX <- FRAX[-c(1)]
names(FRAX)<- c("GENDER", "RACE", "AGE","CRF","BMI","Risk", "Range_lower", "Range_upper")

Diagnosis <- ("H:/Hip fracture/Diagnosis_acriscit.xlsx")
History <- ("H:/Hip fracture/History_acriscit.xlsx")
Lab <- ("H:/Hip fracture/Lab_acriscit.xlsx")
Meds <- ("H:/Hip fracture/Meds_acriscit.xlsx")
Patients <- ("H:/Hip fracture/Patients_acriscit.xlsx")
Procedures <- ("H:/Hip fracture/Procedures_acriscit.xlsx")
Vitals <- ("H:/Hip fracture/Vitals_acriscit.xlsx")

#Function to import excel data with multiple tabs as a list
read_excel_allsheets <- function(filename, tibble = TRUE) {
  tab_names <- excel_sheets(path = filename)
  #Below is the slow line
  filename <- map(tab_names, ~read_excel(path = filename, sheet = .x, col_types = "text"))
  filename <- filename %>% set_names(tab_names) %>% bind_rows ()
}

#The function above will convert each file below into a list that is easy to work with
filename <- Diagnosis
Diagnosis <- read_excel_allsheets(filename)

filename <- History
History <- read_excel_allsheets(filename)

filename <- Meds
Meds <- read_excel_allsheets(filename)

filename <- Patients
Patients <- read_excel_allsheets(filename)

filename <- Procedures
Procedures <- read_excel_allsheets(filename)

filename <- Vitals
Vitals <- read_excel_allsheets(filename)
  
filename <- Lab
Lab <- read_excel_allsheets(filename)


#In this chunk we are organizing data so that relevant information can be more easily collected
#- All the documents we imported above will be combined into one large tibble
#- We will create an empty tibble that will be filled with info to calculate FRAX score at the time of fracture for each patient

#Accumulate data tibbles as a list
tibble_names <- list(
  Vitals, 
  Meds, 
  Lab, 
  Procedures, 
  Diagnosis,
  History
  )
#Title each list element appropriately
tibble_titles <- c(
  "Vitals", 
  "Meds", 
  "Lab", 
  "Procedures",
  "Diagnosis",
  "History"
  )
names(tibble_names) <- tibble_titles

#rename columns and remove spaces from column names
col_names <- c(
  "ID", 
  "VISIT", 
  "DATE",
  "CODE",
  "DESC", 
  "TEXT_VALUE", 
  "NUMERIC_VALUE", 
  "UNITS",
  "MODIFIER"
)
#This loop applies the new column names to each list element in 'tibble_names' and converts dates to mdy_hms
for (i in 1:length(tibble_names)){
  names(tibble_names[[i]]) <- col_names[1:ncol(tibble_names[[i]])]
  tibble_names[[i]]$DATE <- mdy_hm(
    tibble_names[[i]]$DATE, 
    quiet = TRUE
    )
}
#Flatten list into a tibble
x <- tibble_names %>%
  bind_rows()

#remove the xl files. Note that if you want to search one of these files rather than x (because it is faster) you can search tibble_names$...
rm(Diagnosis)
rm(Lab)
rm(Meds)
rm(Patients)
rm(Vitals)
rm(History)

#Do the same process for demographic data
col_names <- c(
  "ID", 
  "AGE_NOW", 
  "AGE_AT_DEATH",
  "BIRTH_DATE",
  "DEATH_DATE",
  "MATCH",
  "GENDER", 
  "RACE", 
  "MIXED RACE", 
  "ETHNICITY"
)
names(Patients) <- col_names[1:ncol(Patients)]

  Patients <- Patients %>%
    mutate(BIRTH_DATE = mdy_hm(BIRTH_DATE))
  
 Patients <- Patients %>%
    mutate(DEATH_DATE = mdy_hm(DEATH_DATE))
y <- Patients


#Now that the patient data is more readable, we will create a tibble that contains data needed to calculate a FRAX score for each patient. This will primarily be accomplished with the use of a loop.

#The below lines will select patients with fracture then remove instances with traumatic fracture or fractures of the hands, feet, face or skull (non major osteoporotic fractures). It is assumed that all other fractures in this >50 y/o population are major osteoporotic fractures. Like the below lines it will then group fracture events by patient ID, arrange them by date and exclude all but the first fracture event using 'distinct ()'
#Note that this does not exclude patients who have had a non-MOF but excludes that diagnosis from analysis

ID_fracs <- tibble_names$Diagnosis %>%
  filter(grepl('[Ff]racture', DESC)) %>%
  #Exclude S62 hand, S92, S99 foot, M88 Osteodeformans/Paget's disease of bone, S02 Face/skull, V trauma due to a collision, Y04 assault by bodily force, C79.5 metastatic disease to bone
  filter(!grepl('S62|S92|S99|M88|S02|V|Y04|C79.5', CODE)) %>%
  #Remove diagnoses from the problem list
  filter(!grepl('PROBLEM:ACTIVE', TEXT_VALUE)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE) %>% 
  distinct(ID, .keep_all = TRUE)
#Calculate the number of patients excluded by this step. Both 'y' and 'ID_fracs' have one row for each patient
number_excluded_frac <- nrow(y) - nrow(ID_fracs)

#The below line will select patients with hip fracture
Hip_fracs <- tibble_names$Diagnosis %>%
  #filter for ICD code M84.359, M84.459, M84.659 which indicates a hip fracture
  filter((grepl('M84.359|M84.459|M84.659|820|S72.0|S72.90XA|S72.91XA|S72.92XA', CODE))) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE) %>% 
  distinct(ID, .keep_all = TRUE)
#Calculate the number of patients excluded by this step
number_excluded_hip <- nrow(y) - nrow(Hip_fracs)

#The below chunk will create a tibble for each FRAX risk factor that includes all patients. For example 'Smoker' is a tibble that contains smoking information on all patients with a history of smoking. An individual patient will be selected from this tibble within the loop. This was done to increase the efficiency of the loop by minimizing the number of functions within the loop


# alternative to 'filter(grepl())' is library(stringr), x %>% filter(str_detect(rowname, "^ICD-10 code")). '^' means starts with

Fracture <- x %>% 
     filter(grepl('[Ff]racture', DESC)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)

#Smoker at the time of the fracture
Smoker <- x %>% 
     filter(grepl('[Ss]moker', DESC) & grepl('[Cc]urrent', DESC)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)

#A diagnosis of alcohol use (ICD-10 codes containing F10 or corresponding ICD-9 codes (303 alcohol dependance, 291 alcohol induced mental disorder, 3050 alcohol abuse, K70 alcoholic liver disease), which encompasses alcohol abuse, alcohol dependence and alcohol use, unspecified) is used as a surrogate for "3 or more units of alcohol per day" as specified by FRAX
EToH <- x %>% 
     filter(grepl('F10|3050|291|303|V11.3|K70', CODE)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)

#RA diagnosis of Rheumatoid Arthritis 
RA <- x %>% 
     filter(grepl('M05|M06|714', CODE)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)

#According to FRAX this includes osteoporosis (M81, 7330) and diseases strongly associated with osteoporosis including type I (insulin dependent) diabetes (E10, 250), osteogenesis imperfecta in adults (Q78.0, 75651), untreated long-standing hyperthyroidism E05, 242)(these are the codes for hyperthyroidism irrespective of treatment), hypogonadism (E29.1, 2572 (male), E28.3, 25631, 25639 (female) including) premature menopause (<45 years), chronic malnutrition (intestinal malabsorption K90, 579, nutritional malabsorption E41-46, 261-263, 579), or malabsorption and chronic liver disease (571-573, K71-K76)
#An alternative to malabsorption could be to consider BMI <15
Osteoporosis <- x %>% 
     filter(grepl('M81|7330|E10|250|Q78.0|75651|E05|242|E29.1|2572|E28.3|25631|25639|k90|579|E41|E42|E43|E44|E45|E46|261|262|263|579|571|572|573|K71|K72|K73|K74|K75|K76', CODE)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)

#FRAX considers parental hip fracture a risk. We have substituted parental osteoporosis for this risk factor because there is a searchable diagnostic code
Parent_Osteoporosis <- x %>% 
     filter(grepl('Z82.62|V17.81', CODE)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)

#Current exposure to oral glucocorticoids or exposure to oral glucocorticoids for more than 3 months at a dose eqivalent to prednisolone > 5mg daily
Glucocorticoid <- tibble_names$Meds %>% 
     filter(MODIFIER == 'mg' &
              grepl('[Bb]eclomethasone|[Bb]etamethasone|[Bb]udesonide|[Cc]ortisone|[Dd]examethasone|[Hh]ydrocortisone|[Mm]ethylprednisolone|[Pp]rednisolone|[Pp]rednisone|[Tt]riamcinolone', DESC)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)

```
The below loop will collect and distill information relevant to calculating FRAX score FOR PATIENTS WHO HAVE HAD A FRACTURE ONLY. Each cycle of this loop will collect data for 1 patient and will populate the tibble 'Fracture_Dates'
```{r loop}
#Create an empty tibble ('Fracture_Dates') with the below columns. It will later be populated by each cycle of the loop
Fracture_Dates <- tibble(
  ID = numeric(),
  DATE = mdy_hms(),
  CODE = character(),
  DESC = character(),
  HT.in = numeric(), 
  WT.in = numeric(),
  AGE = numeric(),
  Birth_Date = mdy_hms(),
  ageatfracture = numeric(),
  ageatdeath = numeric(),
  GENDER = character(), 
  Og_RACE = character(), 
  CRF = numeric(),
  #CRF is Combined Risk Factors
  BMI = numeric()
)

#Tibble where the risk factors will be compiled
CRF <- tibble(
  ID = numeric(),
  Fracture = character(),
  Smoker = character(),
  EToH = character(),
  RA = character(),
  Osteoporosis = character(),
  Parent_Osteoporosis = character(),
  Glucocorticoid_rf = character()
)
#Collect risk factors into above tibble. If there is no observation prior to DATE, it is assumed that the risk factor was not present. For example, if there is no prior fracture documented, it is assumed that there was no prior fracture

for (i in 1:nrow(ID_fracs)){
  id <- as.double(ID_fracs$ID[i])

loop_ID1 <- ID_fracs %>% 
    filter(ID == id)
#Below are common variables that will be used throughout the loop
date <- loop_ID1$DATE[1]
CODE <- loop_ID1$CODE[1]
DESC <- loop_ID1$DESC[1]

#REMOVE THIS IF YOU ARE NOT LOOKING FOR HIP FRACTURES ONLY!!!
#####

#Create a tibble with one patients into
loop_ID <- x %>% 
    filter(ID == id)
#Select only the info prior to fracture
loop_ID_prefrac <- loop_ID %>% filter(DATE <= date)

#Height and weight closest to the time of fracture
HT <- (loop_ID %>% 
         filter(CODE == "HT") %>% 
         arrange(desc(DATE)) %>% 
         select(NUMERIC_VALUE)
       )[1,1]
WT <- (loop_ID_prefrac %>% 
         filter(CODE == "WT") %>% 
         arrange(desc(DATE)) %>% 
         select(NUMERIC_VALUE)
       )[1,1]

#If there is no height or weight prior to the fracture date, use the oldest measurement
HT <- as.double(ifelse(is.na(HT), ((
  loop_ID %>% 
    filter(CODE == "HT") %>% 
    arrange(desc(DATE)) %>% 
    select(NUMERIC_VALUE))[1,1]),
  HT))
WT <- as.double(ifelse (is.na(WT), ((
  loop_ID %>% 
    filter(CODE == "WT") %>% 
    arrange(desc(DATE)) %>% 
    select(NUMERIC_VALUE))[1,1]), 
  WT))

#Calculate BMI
BMI <- (703*as.double(WT)/as.double(HT)^2)

yID <- y %>% filter(ID == id)
#Calcuate age at time of fracture
Birth_Date <- (yID %>% select(BIRTH_DATE))[[1]]
ageatfracture <- (date - Birth_Date) %>% as.double/365.25
ageatdeath <- as.double(yID %>% select(AGE_AT_DEATH))

#Select gender
Gender <- yID %>% select(GENDER)

#Select race
Race <- yID %>% select(RACE)

# alternative to 'filter(grepl())' is library(stringr), x %>% filter(str_detect(rowname, "^ICD-10 code")). '^' means starts with

Fracture_loop <- as.character(
  (filter(Fracture, id == ID & DATE <= date) %>%
  select(DESC))[1,2]
  )
#Smoking within the last 2 years
Smoker_loop <- as.character(
  (filter(Smoker, id == ID & DATE <= date & DATE >= date-years(2)) %>% 
     select(DESC))[1,2]
  )
#A diagnosis of alcohol use (ICD-10 codes containing F10 or corresponding ICD-9 codes (303 alcohol dependance, 291 alcohol induced mental disorder, 3050 alcohol abuse), which encompasses alcohol abuse, alcohol dependence and alcohol use, unspecified) is used as a surrogate for "3 or more units of alcohol per day" as specified by FRAX
EToH_loop <- as.character(
  (filter(EToH, id == ID & DATE <= date) %>%
     select(CODE))[1,2]
  )
#A diagnosis of Rheumatoid Arthritis 
RA_loop <- as.character(
  (filter(RA, id == ID & DATE <= date) %>%
     select(CODE))[1,2]
  )
#According to FRAX this includes osteoporosis (M81, 7330) and diseases strongly associated with osteoporosis including type I (insulin dependent) diabetes (E10, 250), osteogenesis imperfecta in adults (Q78.0, 75651), untreated long-standing hyperthyroidism E05, 242)(these are the codes for hyperthyroidism irrespective of treatment), hypogonadism (E29.1, 2572 (male), E28.3, 25631, 25639 (female) including) premature menopause (<45 years), chronic malnutrition (intestinal malabsorption K90, 579, nutritional malabsorption E41-46, 261-263), or malabsorption and chronic liver disease
#An alternative to malabsorption could be to consider BMI <15
Osteoporosis_loop <- as.character(
  (filter(Osteoporosis, id == ID & DATE <= date) %>% 
     select(CODE))[1,2]
  )
#FRAX considers parental hip fracture a risk. We have substituted parental osteoporosis for this risk factor because there is a searchable diagnostic code
Parent_Osteoporosis_loop <- as.character(
  (filter(Parent_Osteoporosis, id == ID) %>% 
     select(CODE))[1,2]
  )
#Current exposure to oral glucocorticoids or exposure to oral glucocorticoids for more than 3 months at a dose eqivalent to prednisolone > 5mg daily
Glucocorticoid_data <- (
  filter(Glucocorticoid, id == ID & DATE <= date)
  )
Glucocorticoid_rf <- ifelse(is.na(Glucocorticoid_data$ID[1]), NA, "TRUE")
#Exclude glucocorticoids taken for less than 3 months


#Add each risk factor to the 'CRF' tibble
CRF <- CRF %>% add_row(
ID = id,
Fracture = Fracture_loop,
Smoker = Smoker_loop,
EToH = EToH_loop,
RA = RA_loop,
Osteoporosis = Osteoporosis_loop,
Parent_Osteoporosis = Parent_Osteoporosis_loop,
Glucocorticoid_rf = as.character(Glucocorticoid_rf)
)
#Change the name of the tibble to calculate CRF score below so as to preserve the CRF tibble for retrospective analysis
CRF_calc <- CRF %>% filter(ID == id) %>% select(!ID)
#Convert risk factors to 0 (not present) or 1 (present)
CRF_calc[!is.na(CRF_calc)] <- '1'
CRF_calc[is.na(CRF_calc)] <- '0'
#Calculate CRF
CRF_score <- sum(as.double(CRF_calc))

#Put all the data you collected into 'Fracture_Dates'
Fracture_Dates = Fracture_Dates %>% add_row(
ID = id,
DATE = date,
CODE = CODE,
DESC = DESC,
HT.in = as.double(HT), 
WT.in = as.double(WT), 
AGE = as.double(ageatfracture), 
Birth_Date = Birth_Date,
ageatfracture = as.double(ageatfracture),
ageatdeath = as.double(ageatdeath),
GENDER = as.character(Gender), 
Og_RACE = as.character(Race), 
CRF = as.double(CRF_score),
BMI = as.double(BMI)
)
rm(loop_ID)
rm(loop_ID1)
rm(loop_ID_prefrac)
rm(yID)
}

#Save the original Fracture_Dates because the below chunks will make changes to it
Fracture_Dates_Og <- Fracture_Dates

#add time from fracture to death via mutate()
Fracture_Dates <- Fracture_Dates %>% mutate(time_to_death = ageatdeath - ageatfracture)
#remove anyone < 50 years old at the first fracture. This should not include anyone in the cohort, but will catch anyone extra
Fracture_Dates <- Fracture_Dates %>% filter(ageatfracture >= 50)
Fracture_Dates$ID <- as.double(Fracture_Dates$ID)

#Calculate FRAX score below

#add death date to Fracture_Dates and change the name of the tibble to 'px'
y$ID <- as.double(y$ID)
px <- y %>%
  select(ID, DEATH_DATE, AGE_AT_DEATH) %>%
  right_join(Fracture_Dates, by = 'ID')

#Round BMI down to the nearest 5
floorfive <- function(x) {floor(x/5)*5}
px <- px %>% mutate(
  BMI = ifelse((BMI<45)&(BMI>15),floorfive(BMI), 
               ifelse (BMI>45,45,15)
               )
  )

#Round age down to the nearest 5
px <- px %>% mutate (
  AGE = ifelse((AGE<90)&(AGE>50),floorfive(AGE), 
               ifelse (AGE<50,50,90)
               )
  )

#Change race to white if not white, black, hispanic or asian
#Original race will be maintained the the column 'Og_RACE'
px$Og_RACE <- as.character(px$Og_RACE)
Races <- c("W","B","H","A")
px <- px %>% mutate(RACE = ifelse(
  (Og_RACE %in% Races), Og_RACE, "W")
  )

#Join FRAX and px to calculate 10 year risk of major osteoporotic fracture (do you mean major hip fracture?)
fx <- px %>% 
  left_join(FRAX, by= c("GENDER","RACE","AGE","CRF","BMI"), copy=FALSE, keep=FALSE)%>%
  select(ID, Risk, Range_upper, Range_lower) %>% 
  arrange(ID) %>%
  na.omit()

Fracture_Dates_fx <- Fracture_Dates %>% left_join(fx, by = 'ID')

write.csv(Fracture_Dates_fx, "C:\\Users\\acriscit\\Documents\\Fracture_Dates_fx.csv")
write.csv(CRF, "C:\\Users\\acriscit\\Documents\\CRF_fx.csv")
```
#Descriptors
```{r descriptors}

percent_complete_cases <- (nrow(fx)/nrow(Fracture_Dates))*100
percent_missing_FRAX <- ((nrow(fx)-nrow(Fracture_Dates_fx))/percent_complete_cases)*100
percent_missing_HT <- (nrow(is.na(Fracture_Dates_fx$HT.in))/nrow(Fracture_Dates_fx))*100
percent_missing_WT <- (nrow(is.na(Fracture_Dates_fx$WT.in))/nrow(Fracture_Dates_fx))*100

Flow <- tibble(
  percent_complete_cases = numeric(),
  number_excluded_frac = numeric(),
  number_excluded_hip = numeric(),
  percent_missing_FRAX = numeric(),
  percent_missing_WT = numeric(),
  percent_missing_HT = numeric()
)

Flow <- Flow %>% add_row(
  percent_complete_cases = as.double(percent_complete_cases),
  number_excluded_frac = as.double(number_excluded_frac),
  number_excluded_hip = as.double(number_excluded_frac),
  percent_missing_FRAX = as.double(percent_missing_FRAX),
  percent_missing_WT = as.double(percent_missing_WT),
  percent_missing_HT = as.double(percent_missing_HT)
)
write.csv(Flow, "C:\\Users\\acriscit\\Documents\\Flow.csv")

rm(px)
rm(fx)
rm(Fracture_Dates)
```
#FOR HIP FRACTURE ONLY
#Ensure that patients from Fracture_Dates_fx actually had a hip fracture first
- Vertebral: 805|806|733.13|S12|M48.50|M80.08
- Humerus (distal and shaft): 812.01|812.23|733.11|S42.2|M84.429A
- Forearm: 813.45|813.8|733.12|813.0|S52|M84.429A
```{r hip}
#Collect the ID's of patients with MOF. MOF_ID contains the date of MOF, Dx code and patient ID
MOF_ID <- tibble_names$Diagnosis %>% 
     filter(grepl('805|806|733.13|S12|M48.50|M80.08|812.01|812.23|733.11|S42.2|M84.429A|813.45|813.8|733.12|813.0|S52|M84.429A', CODE)) %>%
  group_by(ID) %>%
  arrange(DATE, .by_group = TRUE) %>%
  select(ID, DATE, CODE)
```
The tibble 'Fracture_Dates' now contains a row for each patient and associated information necessary to calculate FRAX score. Below we will use the 'FRAX' tibble create from publicly available FRAX paper charts (https://www.sheffield.ac.uk/FRAX/charts.aspx) to calculate the FRAX score for each patient.

Discern whether the patients with fracture have diabetes.
What is the duration of disease prior to fracture?
Were they well controlled?
```{r Diabetes calculator}
#Select patients with type 2 DM using ICD-10/9 code 'E11' or '250'
All_DM <- x %>% 
  filter((grepl('E11|250', CODE))) %>% 
  #Select the ID and Date of first DM diagnosis in the record
  select(ID, DATE) %>% group_by(ID) %>%
  arrange(DATE, .by_group = TRUE) %>%
  distinct(ID, .keep_all = TRUE) %>%
  transform(ID = as.double(ID))
#Select patients with DM for whom we have a Frax score
fx <- Fracture_Dates_fx %>% filter(!is.na(Risk))
ID_DM <- inner_join(All_DM, fx, by = 'ID')
rm(All_DM)

#calculate the time different between fracture and diagnosis of DM. A negative value indicates that the diagnosis was made after this visit
ID_DM <- ID_DM %>% mutate(
  days_before_fracture = difftime(DATE.y, DATE.x, units = "days"))
ID_DM <- ID_DM %>% mutate(
  DM_Dx_postvisit = ifelse(days_before_fracture < 0, "TRUE", "FALSE"))

#Pre-loop
Insulin <- x %>%
  filter(grepl('[Ii]nsulin', DESC)) %>%
  group_by(ID) %>%
  arrange(desc(DATE), .group_by = TRUE)
BG <- x %>%
  filter(grepl('GLU', CODE)) %>%
  filter(grepl('mg/dl|MG/DL', UNITS)) %>%
  filter(NUMERIC_VALUE != 0 & !is.na(NUMERIC_VALUE)) %>%
  arrange(desc(DATE))
HBA1C <- x %>%
  filter(grepl('HBA1C', CODE)) %>%
  filter(UNITS == '%') %>%
  arrange(desc(DATE))
```
Diabetes loop
```{r DM loop}
#Tibble to be filled by loop
DM_dates_fx <- tibble (
  ID = numeric(),
  HBA1C_max = numeric(),
  HBA1C_max_date = character(),
  HBA1C_avg = numeric(),
  Insulin_start = character(),
  long_term_insulin = character(),
  BG = numeric(),
  BG_date = character(),
  BG1yavg = numeric(),
  BG1yr_num_values = numeric()
)

for(i in 1:nrow(ID_DM)) {
  id <- as.double(ID_DM$ID[i])
  
  #Date of patients first fracture
  fracture_date <- ID_DM$DATE.y[i]

 # pt_DM <- ID_DM$DATE.x[i]

#HBA1C's prior to fracture and date of measurement
HBA1C_loop <- HBA1C %>% 
  filter(ID == id) %>%
  filter(DATE <= fracture_date)

HBA1C_max <- as.double(max(HBA1C_loop$NUMERIC_VALUE))
#HBA1C_max <- ifelse(is.na(HBA1C_max),0,HBA1C_max)

HBA1C_max_date <- (HBA1C_loop %>%
                     filter(NUMERIC_VALUE == HBA1C_max))$DATE[1]

#Average HBA1C in the year before fracture
HBA1C_avg <- HBA1C_loop$NUMERIC_VALUE %>% as.double() %>% mean()
#HBA1C_avg <- ifelse(is.na(HBA1C_avg),0,HBA1C_avg)

#Insulin and BG
Insulin_dates <- Insulin %>%
  filter(id == ID) %>%
  filter(DATE < fracture_date)
Insulin_start <- Insulin_dates$DATE[1]
#Long term insulin use noted in problem list
long_term_insulin <- (Insulin_dates %>%
  filter(grepl('[Ll]ong term|[Ll]ong-term', DESC)))$DESC[1]
  
BGgen <- BG %>%
  filter(ID == id) %>%
  filter(DATE < fracture_date)
BG_date <- BGgen$DATE[1]
BGnow <- as.double(BGgen$NUMERIC_VALUE)[1]
#BG averaged over 1 year prior to fracture if there are 2 or more values recorded
BG1yr <- BGgen %>% filter(DATE >= fracture_date - years(1))
BG1yr_num_values <- BG1yr %>% nrow()
BG1yavg <- ifelse(BG1yr_num_values > 1, BG1yr$NUMERIC_VALUE %>% as.double() %>% mean(), NA)

rm(BG1yr)

  DM_dates_fx <- DM_dates_fx %>% add_row(
   ID = id,
   HBA1C_max = as.double(HBA1C_max),
   HBA1C_max_date = as.character(HBA1C_max_date),
   HBA1C_avg = as.double(HBA1C_avg),
   Insulin_start = as.character(Insulin_start),
   long_term_insulin = long_term_insulin,
   BG = BGnow,
   BG_date = as.character(BG_date),
   BG1yavg = BG1yavg,
   BG1yr_num_values = BG1yr_num_values
)
  rm(HBA1C_loop)
  rm(Insulin_dates)
  rm(BGgen)
  rm(BG1yr)
}

#replace 'NaN' with NA
DM_dates_fx$HBA1C_avg[DM_dates_fx$HBA1C_avg == "NaN"] <- NA
#Make long term insulin true or false
DM_dates_fx$long_term_insulin <- ifelse(!is.na(DM_dates_fx$long_term_insulin), "TRUE", "FALSE")
#Join the info pulled into 'ID_DM' form FRAX loop
DM_dates_fx <- DM_dates_fx %>%
  left_join(ID_DM, by='ID')
#Rename the date columns
DM_dates_fx <- DM_dates_fx %>% rename(fx_DATE = DATE.y, DM_DATE = DATE.x)

write.csv(DM_dates_fx, "C:\\Users\\acriscit\\Documents\\DM_dates_fx.csv")

rm(ID_DM)
rm(BG)
rm(Insulin)
rm(HBA1C)
```
DXA info
``` {r DXA}
Fracture_Dates_fx$ID <- as.character(Fracture_Dates_fx$ID)

DXA <- tibble_names$Procedures %>%
  filter(grepl('DXA', DESC)) %>%
  select(ID, DESC, DATE) %>%
  mutate(DXA = "TRUE")

write.csv(DXA, "C:\\Users\\acriscit\\Documents\\DXA.csv")
```
Visit info
```{r Visits}
Visits <- ("H:/Hip fracture/Visits_acriscit.xlsx")

filename <- Visits
Visits <- read_excel_allsheets(filename)

col_names <- c(
  "ID", 
  "VISIT", 
  "START_DATE",
  "END_DATE",
  "AGE_AT_VISIT", 
  "ADMIT_SOURCE", 
  "ADMIT_TYPE", 
  "LOS",
  "TYPE",
  "ENCOUTER_TYPE",
  "CCI"
)
#Applies the new column names to each list element in 'tibble_names' and converts dates to mdy_hms
  names(Visits) <- col_names[1:ncol(Visits)]
  
  #Convert visits dates to dates and use as.Date to remove the time element
  Visits <- Visits %>% mutate(START_DATE = as.Date(mdy_hm(START_DATE)))
  Visits <- Visits %>% mutate(END_DATE = as.Date(mdy_hm(END_DATE)))

  #One visit is often tracked as multiple visits on the same date. Remove duplicate visits
  Visits <- Visits %>%
    distinct(ID, START_DATE, .keep_all = TRUE) %>% group_by(ID) %>% arrange(START_DATE, .by_group = TRUE)
  
```
run the loop below
```{r visit_loop}

#Information about number of visits and date of visits relative to fracture date
Visit_Dates <-  tibble(
  ID = numeric(),
  first_visit_date = mdy(),
  last_visit_date = mdy(),
  num_visits_before_frac = numeric(),
  num_visits_after_frac = numeric(),
  CCI = numeric ()
)

for (i in 1:nrow(Fracture_Dates_fx)){
  id <- as.double(Fracture_Dates_fx$ID[i])

date <- (Fracture_Dates_fx %>% 
    filter(ID == id) %>%
    select(ID, DATE))[[1,2]]

#Create a tibble with one patients into
visit <- Visits %>% 
    filter(ID == id)

first_visit_date <- visit$START_DATE[1]
last_visit_date <- (
  visit %>% 
    arrange(desc(START_DATE))
  )$START_DATE[1]

num_visits_before_frac <- visit %>% 
  filter(START_DATE < date) %>%
  #mutate(START_DATE = floor_date(START_DATE, "day")) %>%
  #group_by(START_DATE) %>% 
  #distinct(START_DATE, .keep_all = TRUE) %>%
  nrow()
num_visits_after_frac <- visit %>% 
  filter(START_DATE > date) %>% 
  #mutate(START_DATE = floor_date(START_DATE, "day")) %>%
  #group_by(START_DATE) %>% 
  #distinct(START_DATE, .keep_all = TRUE) %>%
  nrow()
#LOS_fx <- visit %>% 
  #filter(TYPE == 'I|S') %>% 
 # filter(START_DATE <= date & END_DATE >= date) %>%
 # select(LOS) %>% as.double()[[1]]

#Finds the most recent CCI prior to the fracture date
CCI <- visit %>%
  filter(!is.na(CCI) & START_DATE < date) %>%
 # filter(START_DATE <= date) %>%
  arrange(desc(START_DATE)) %>%
  distinct(ID, .keep_all = TRUE) %>%
  select(CCI)
CCI <- as.double(CCI[1,2])

Visit_Dates <- Visit_Dates %>% add_row(
  ID = id,
  first_visit_date = first_visit_date,
  last_visit_date = last_visit_date,
  num_visits_before_frac = num_visits_before_frac,
  num_visits_after_frac = num_visits_after_frac,
  CCI = as.double(CCI)
)
rm(visit)
}

write.csv(Visit_Dates, "C:\\Users\\acriscit\\Documents\\Visit_Dates.csv")
rm(Visit_Dates)
#Join Visits to tibble_names and save tibble_names as a patient database for future use
#remove 'test' once you've tested this
test <- list.append(tibble_names, Visits, after = Vitals)
rm(Visits)
```
Osteo meds
```{r osteo meds}
#FDA approved osteoporosis medication usage
osteo_med_all <- x %>% 
  filter(
    grepl('[Ee]tidronate|[Aa]lendronate|[Ii]bandronate|[Zz]oledronic|[Rr]isedronate|[Cc]alcitonin|[Dd]enosumab|[Ee]strogen|[Rr]aloxifene|[Bb]azodoxifene|[Rr]omosozumab|[Tt]eriparatide|[Aa]baloparatide', DESC) &
    !grepl('[Rr]eceptor', DESC)
         ) %>%
  filter(MODIFIER == "mg") %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)
```
Osteo loop
```{r osteo_loop}
Osteo_Meds <- tibble(
  ID = numeric(),
  osteo_med_start_date = character(),
  osteo_med_start_dose_mg = numeric(),
  osteo_med_last_date = character(),
  osteo_med_name = character()
)

for (i in 1:nrow(Fracture_Dates_fx)){
  id <- as.double(Fracture_Dates_fx$ID[i])

date <- (Fracture_Dates_fx %>% 
    filter(ID == id) %>%
    select(ID, DATE))[[1,2]]
  
  #FDA approved osteoporosis medication usage
osteo_med <- (filter(osteo_med_all, id == ID))
osteo_med_name <- as.character(osteo_med$DESC[1])
osteo_med_start_date <- osteo_med$DATE[1]
osteo_med_start_dose <- osteo_med$UNITS[1]
osteo_med_last_date <- (osteo_med %>%
  arrange(desc(DATE)))$DATE[1]

Osteo_Meds <- Osteo_Meds %>% add_row(
  ID = id,
  osteo_med_start_date = as.character(osteo_med_start_date),
  osteo_med_start_dose_mg = as.double(osteo_med_start_dose),
  osteo_med_last_date = as.character(osteo_med_last_date),
  osteo_med_name = as.character(osteo_med_name)
)
rm(osteo_med)
}

Osteo_Meds <- na.omit(Osteo_Meds)
write.csv(Osteo_Meds, "C:\\Users\\acriscit\\Documents\\Osteo_Meds.csv")
rm(Osteo_Meds)
```
FRAX scores 5 years prior to fracture
```{r 5 years before}
Fracture_Dates_5 <- tibble(
  ID = numeric(),
  DATE = mdy_hms(),
  CODE = character(),
  DESC = character(),
  HT.in = numeric(), 
  WT.in = numeric(),
  AGE = numeric(),
  Birth_Date = mdy_hms(),
  ageatfracture = numeric(),
  ageatdeath = numeric(),
  GENDER = character(), 
  Og_RACE = character(), 
  CRF = numeric(),
  #CRF is Combined Risk Factors
  BMI = numeric()
)

#Tibble where the risk factors will be compiled
CRF <- tibble(
  ID = numeric(),
  Fracture = character(),
  Smoker = character(),
  EToH = character(),
  RA = character(),
  Osteoporosis = character(),
  Parent_Osteoporosis = character(),
  Glucocorticoid_rf = character()
)

#ID_fracs_5 and Hip_fracs_5 will set the date as 5 years prior to fracture
ID_fracs_5 <- tibble_names$Diagnosis %>%
  filter(grepl('[Ff]racture', DESC)) %>%
  #Exclude S62 hand, S92, S99 foot, M88 Osteodeformans/Paget's disease of bone, S02 Face/skull, V trauma due to a collision, Y04 assault by bodily force, C79.5 metastatic disease to bone
  filter(!grepl('S62|S92|S99|M88|S02|V|Y04|C79.5', CODE)) %>%
  #Remove diagnoses from the problem list
  filter(!grepl('PROBLEM:ACTIVE', TEXT_VALUE)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE) %>% 
  distinct(ID, .keep_all = TRUE) %>%
  mutate(DATE = DATE - years(5))

#The below line will select patients with hip fracture
Hip_fracs_5 <- tibble_names$Diagnosis %>%
  #filter for ICD code M84.359, M84.459, M84.659 which indicates a hip fracture
  filter((grepl('M84.359|M84.459|M84.659|820|S72.0|S72.90XA|S72.91XA|S72.92XA', CODE))) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE) %>% 
  distinct(ID, .keep_all = TRUE) %>%
  mutate(DATE = DATE - years(5))
```
#FOR HIP FRACTURE ONLY
#remove patients with a history of major osteoporotic fracture prior to the hip fracture
- Vertebral: 805|806|733.13|S12|M48.50|M80.08
- Humerus (distal and shaft): 812.01|812.23|733.11|S42.2|M84.429A
- Forearm: 813.45|813.8|733.12|813.0|S52|M84.429A
```{r hip}
#Collect the ID's of patients with MOF. MOF_ID contains the date of MOF, Dx code and patient ID
MOF_ID <- x %>% 
     filter(grepl('805|806|733.13|S12|M48.50|M80.08|812.01|812.23|733.11|S42.2|M84.429A|813.45|813.8|733.12|813.0|S52|M84.429A', CODE)) %>%
  group_by(ID) %>%
  arrange(DATE, .by_group = TRUE) %>%
  select(ID, DATE, CODE)

```
The below loop will collect and distill information relevant to calculating FRAX score FOR PATIENTS WHO HAVE HAD A FRACTURE ONLY. Each cycle of this loop will collect data for 1 patient and will populate the tibble 'Fracture_Dates_5'
```{r loop}
for (i in 1:nrow(ID_fracs_5)){
  id <- as.double(ID_fracs_5$ID[i])

loop_ID1 <- ID_fracs_5 %>% 
    filter(ID == id)
#Below are common variables that will be used throughout the loop
date <- loop_ID1$DATE[1]
CODE <- loop_ID1$CODE[1]
DESC <- loop_ID1$DESC[1]

#REMOVE THIS IF YOU ARE NOT LOOKING FOR HIP FRACTURES ONLY!!!
#####

#Create a tibble with one patients into
loop_ID <- x %>% 
    filter(ID == id)
#Select only the info prior to fracture
loop_ID_prefrac <- loop_ID %>% filter(DATE <= date)

#Height and weight closest to the time of fracture
HT <- (loop_ID %>% 
         filter(CODE == "HT") %>% 
         arrange(desc(DATE)) %>% 
         select(NUMERIC_VALUE)
       )[1,1]
WT <- (loop_ID_prefrac %>% 
         filter(CODE == "WT") %>% 
         arrange(desc(DATE)) %>% 
         select(NUMERIC_VALUE)
       )[1,1]

#If there is no height or weight prior to the fracture date, use the oldest measurement
HT <- as.double(ifelse(is.na(HT), ((
  loop_ID %>% 
    filter(CODE == "HT") %>% 
    arrange(desc(DATE)) %>% 
    select(NUMERIC_VALUE))[1,1]),
  HT))
WT <- as.double(ifelse (is.na(WT), ((
  loop_ID %>% 
    filter(CODE == "WT") %>% 
    arrange(desc(DATE)) %>% 
    select(NUMERIC_VALUE))[1,1]), 
  WT))

#Calculate BMI
BMI <- (703*as.double(WT)/as.double(HT)^2)

yID <- y %>% filter(ID == id)
#Calcuate age at time of fracture
Birth_Date <- (yID %>% select(BIRTH_DATE))[[1]]
ageatfracture <- (date - Birth_Date) %>% as.double/365.25
ageatdeath <- as.double(yID %>% select(AGE_AT_DEATH))

#Select gender
Gender <- yID %>% select(GENDER)

#Select race
Race <- yID %>% select(RACE)

# alternative to 'filter(grepl())' is library(stringr), x %>% filter(str_detect(rowname, "^ICD-10 code")). '^' means starts with

Fracture_loop <- as.character(
  (filter(Fracture, id == ID & DATE <= date) %>%
  select(DESC))[1,2]
  )
#Smoking within the last 2 years
Smoker_loop <- as.character(
  (filter(Smoker, id == ID & DATE <= date & DATE >= date-years(2)) %>% 
     select(DESC))[1,2]
  )
#A diagnosis of alcohol use (ICD-10 codes containing F10 or corresponding ICD-9 codes (303 alcohol dependance, 291 alcohol induced mental disorder, 3050 alcohol abuse), which encompasses alcohol abuse, alcohol dependence and alcohol use, unspecified) is used as a surrogate for "3 or more units of alcohol per day" as specified by FRAX
EToH_loop <- as.character(
  (filter(EToH, id == ID & DATE <= date) %>%
     select(CODE))[1,2]
  )
#A diagnosis of Rheumatoid Arthritis 
RA_loop <- as.character(
  (filter(RA, id == ID & DATE <= date) %>%
     select(CODE))[1,2]
  )
#According to FRAX this includes osteoporosis (M81, 7330) and diseases strongly associated with osteoporosis including type I (insulin dependent) diabetes (E10, 250), osteogenesis imperfecta in adults (Q78.0, 75651), untreated long-standing hyperthyroidism E05, 242)(these are the codes for hyperthyroidism irrespective of treatment), hypogonadism (E29.1, 2572 (male), E28.3, 25631, 25639 (female) including) premature menopause (<45 years), chronic malnutrition (intestinal malabsorption K90, 579, nutritional malabsorption E41-46, 261-263), or malabsorption and chronic liver disease
#An alternative to malabsorption could be to consider BMI <15
Osteoporosis_loop <- as.character(
  (filter(Osteoporosis, id == ID & DATE <= date) %>% 
     select(CODE))[1,2]
  )
#FRAX considers parental hip fracture a risk. We have substituted parental osteoporosis for this risk factor because there is a searchable diagnostic code
Parent_Osteoporosis_loop <- as.character(
  (filter(Parent_Osteoporosis, id == ID) %>% 
     select(CODE))[1,2]
  )
#Current exposure to oral glucocorticoids or exposure to oral glucocorticoids for more than 3 months at a dose eqivalent to prednisolone > 5mg daily
Glucocorticoid_data <- (
  filter(Glucocorticoid, id == ID & DATE <= date)
  )
Glucocorticoid_rf <- ifelse(is.na(Glucocorticoid_data$ID[1]), NA, "TRUE")
#Exclude glucocorticoids taken for less than 3 months


#Add each risk factor to the 'CRF' tibble
CRF <- CRF %>% add_row(
ID = id,
Fracture = Fracture_loop,
Smoker = Smoker_loop,
EToH = EToH_loop,
RA = RA_loop,
Osteoporosis = Osteoporosis_loop,
Parent_Osteoporosis = Parent_Osteoporosis_loop,
Glucocorticoid_rf = as.character(Glucocorticoid_rf)
)
#Change the name of the tibble to calculate CRF score below so as to preserve the CRF tibble for retrospective analysis
CRF_calc <- CRF %>% filter(ID == id) %>% select(!ID)
#Convert risk factors to 0 (not present) or 1 (present)
CRF_calc[!is.na(CRF_calc)] <- '1'
CRF_calc[is.na(CRF_calc)] <- '0'
#Calculate CRF
CRF_score <- sum(as.double(CRF_calc))

#Put all the data you collected into 'Fracture_Dates_5'
Fracture_Dates_5 = Fracture_Dates_5 %>% add_row(
ID = id,
DATE = date,
CODE = CODE,
DESC = DESC,
HT.in = as.double(HT), 
WT.in = as.double(WT), 
AGE = as.double(ageatfracture), 
Birth_Date = Birth_Date,
ageatfracture = as.double(ageatfracture),
ageatdeath = as.double(ageatdeath),
GENDER = as.character(Gender), 
Og_RACE = as.character(Race), 
CRF = as.double(CRF_score),
BMI = as.double(BMI)
)
rm(loop_ID)
rm(loop_ID1)
rm(loop_ID_prefrac)
rm(yID)
}

#add time from fracture to death via mutate()
Fracture_Dates_5 <- Fracture_Dates_5 %>% mutate(time_to_death = ageatdeath - ageatfracture)
#remove anyone < 50 years old at the first fracture. This should not include anyone in the cohort, but will catch anyone extra
Fracture_Dates_5 <- Fracture_Dates_5 %>% filter(ageatfracture >= 50)
Fracture_Dates_5$ID <- as.double(Fracture_Dates_5$ID)

#Calculate FRAX score below

#add death date to Fracture_Dates_5 and change the name of the tibble to 'px'
y$ID <- as.double(y$ID)
px <- y %>%
  select(ID, DEATH_DATE, AGE_AT_DEATH) %>%
  right_join(Fracture_Dates_5, by = 'ID')

#Round BMI down to the nearest 5
floorfive <- function(x) {floor(x/5)*5}
px <- px %>% mutate(
  BMI = ifelse((BMI<45)&(BMI>15),floorfive(BMI), 
               ifelse (BMI>45,45,15)
               )
  )

#Round age down to the nearest 5
px <- px %>% mutate (
  AGE = ifelse((AGE<90)&(AGE>50),floorfive(AGE), 
               ifelse (AGE<50,50,90)
               )
  )

#Change race to white if not white, black, hispanic or asian
#Original race will be maintained the the column 'Og_RACE'
px$Og_RACE <- as.character(px$Og_RACE)
Races <- c("W","B","H","A")
px <- px %>% mutate(RACE = ifelse(
  (Og_RACE %in% Races), Og_RACE, "W")
  )

#Join FRAX and px to calculate 10 year risk of major osteoporotic fracture (do you mean major hip fracture?)
fx <- px %>% 
  left_join(FRAX, by= c("GENDER","RACE","AGE","CRF","BMI"), copy=FALSE, keep=FALSE)%>%
  select(ID, Risk, Range_upper, Range_lower) %>% 
  arrange(ID) %>%
  na.omit()

Fracture_Dates_5_fx <- Fracture_Dates_5 %>% left_join(fx, by = 'ID')

write.csv(Fracture_Dates_5_fx, "C:\\Users\\acriscit\\Documents\\Fracture_Dates_5_fx.csv")

rm(px)
rm(fx)
rm(Fracture_Dates_5)
