---
---
title: "FRAX calculator for patients without fracture pulled by month"
author: "Anthony Criscitiello and Ellen Quillen"
date: "9/23/2021"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "FRAX calculator for I2b2 results"
output: html_document
---
```{r setup}
library(tidyverse)
library(readxl)
library(readr)
library(lubridate)
library(tibble)
library(purrr)
library(stringr)
```
^I find it helpful to set everything up at the beggining and you can also make some modifications to how this Rmd turns into an html. Here is cheatsheet on options for R markdown: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
##Introduction
```{r load FRAX ppr charts and patient data}
#load FRAX paper charts (10-year risk of major osteoporotic fracture based on BMI)
FRAX <-  read_csv("H:/Functional Data/Hip fracture/FRAX.csv", col_types = "dccdddddd")
FRAX <- FRAX[-c(1)]
names(FRAX)<- c("GENDER", "RACE", "AGE","CRF","BMI","Risk", "Range_lower", "Range_upper")

#Note that unlike prior data, one excel file contains all information for all patients seen in the ED over the course of one month
thirteen <- ("H:/Health Control (Hip fx)/2013_acriscit.xlsx")

#Function to import excel data with multiple tabs as a list
read_excel_allsheets <- function(filename, tibble = TRUE) {
  tab_names <- excel_sheets(path = filename)
  #Below is the slow line
  filename <- map(tab_names, ~read_excel(path = filename, sheet = .x, col_types = "text"))
  filename <- filename %>% set_names(tab_names) %>% bind_rows ()
}
#The function above will convert each file below into a list that is easy to work with
filename <- thirteen
thirteen <- read_excel_allsheets(filename)
#Remove the 'Patients' part of the tibble by selecting the number of patients/rows
#2789 patients in this group
thirteen_patients <- thirteen[1:2789,1:8]
#Remove these rows and columns from the og tibble
thirteen <- thirteen[-c(1:2789),-c(2:9)]
####pivot longer might be good here, but I would lose the associations between dates, quantities and units
test <- thirteen %>% pivot_longer(cols = 2:27, names_to = TYPE, values_to = VALUE)

##### alternative, try this because the above lines merge the file into one tibble, not a list
test <- "H:/Health Control (Hip fx)/2013_acriscit.xlsx"
#Label each list element with it's associated tab from the excel sheet
tab_names <- excel_sheets(path = test)
#############################################working in github
#Subset the list elements to merge, try to make this into a loop eventually#In this loop, i would be tab names.
tab_names <- tab_names %>% filter(grepl(Meds,***)
#Below is the slow line
test <- map(tab_names, ~read_excel(path = test, sheet = .x))
test <- test %>% set_names(tab_names)

###I want a want to create a list with an element for each data type

col_names <- c(
  "ID", 
  "VISIT", 
  "DATE",
  "CODE",
  "DESC", 
  "TEXT_VALUE", 
  "NUMERIC_VALUE", 
  "UNITS",
  "MODIFIER"
)
for (i in 1:length(test)){
  names(test[[i]]) <- col_names[1:ncol(test[[i]])]
}
#Tidy the data by condensing the 'list' into a 'tibble' using 'bind_rows'
#BEWARE USING X AND Y OR SIMILARLY NON-DESCRIPT VARIABLES FOR OBJECTS YOU NEED TO USE OVER A LONG PERIOD. MORE DESCRIPTIVE NAMES HELP YOU REMEMBER WHAT YOU WERE DOING DOWN THE LINE
x <- bind_rows(test)
x$DATE <- mdy_hm(x$DATE, quiet = TRUE)

```
In this chunk we are organizing data so that relevant information can be more easily collected
- All the documents we imported above will be combined into one large tibble
- We will create an empty tibble that will be filled with info to calculate FRAX score at the time of fracture for each patient
```{r run}
#Accumulate data tibbles as a list
tibble_names <- list(
  Vitals, 
  Meds, 
  Lab, 
  Procedures, 
  Diagnosis,
  History
  )
#Title each list element appropriately
tibble_titles <- c(
  "Vitals", 
  "Meds", 
  "Lab", 
  "Procedures",
  "Diagnosis",
  "History"
  )
names(tibble_names) <- tibble_titles
#rename columns and remove spaces from column names
col_names <- c(
  "ID", 
  "VISIT", 
  "DATE",
  "CODE",
  "DESC", 
  "TEXT_VALUE", 
  "NUMERIC_VALUE", 
  "UNITS",
  "MODIFIER"
)
#This loop applies the new column names to each list element in 'tibble_names' and converts dates to mdy_hms
for (i in 1:length(tibble_names)){
  names(tibble_names[[i]]) <- col_names[1:ncol(tibble_names[[i]])]
  tibble_names[[i]]$DATE <- mdy_hm(
    tibble_names[[i]]$DATE, 
    quiet = TRUE
    )
}
#Flatten list into a tibble
x <- tibble_names %>%
  bind_rows()

#remove the xl files. Note that if you want to search one of these files rather than x (because it is faster) you can search tibble_names$...
rm(Diagnosis)
rm(Lab)
rm(Meds)
rm(Patients)
rm(Vitals)
rm(History)

#Do the same process for demographic data
col_names <- c(
  "ID", 
  "AGE_NOW", 
  "AGE_AT_DEATH",
  "BIRTH_DATE",
  "DEATH_DATE",
  "MATCH",
  "GENDER", 
  "RACE", 
  "MIXED RACE", 
  "ETHNICITY"
)
names(Patients) <- col_names[1:ncol(Patients)]
  Patients <- Patients %>%
    mutate(BIRTH_DATE = mdy_hm(BIRTH_DATE))
  
 Patients <- Patients %>%
    mutate(DEATH_DATE = mdy_hm(DEATH_DATE))
y <- Patients
#Now that the patient data is more readable, we will create a tibble that contains data needed to calculate a FRAX score for each patient. This will primarily be accomplished with the use of a loop.

```
#FROM THIS POINT ON YOU HAVE TWO OPTIONS
1) Upload all files from all months, merge them and remove duplicates
2) Run each file through the loop (faster) and combine after processing to remove duplicates
```{r continue}
```
Create a tibble with a row for each patient with the date of their ED visit and diagnosis at visit
```{r ED}
date_lower <- mdy("5/1/2013")
date_upper <- mdy("5/31/2014")
ID_EM <- tibble_names$Diagnosis %>% filter(TEXT_VALUE == "Primary" & DATE > date_lower & DATE < date_upper) %>% group_by('ID') %>% arrange(DATE, .group_by = TRUE) %>% distinct(ID, .keep_all = TRUE)
#Add a visit filter to this when the data is available
```
#The below chunk will create a tibble for each FRAX risk factor that includes all patients. For example 'Smoker' is a tibble that contains smoking information on all patients with a history of smoking. An individual patient will be selected from this tibble within the loop. This was done to increase the efficiency of the loop by minimizing the number of functions within the loop
```{r risk_factors}
# alternative to 'filter(grepl())' is library(stringr), x %>% filter(str_detect(rowname, "^ICD-10 code")). '^' means starts with
Fracture <- tibble_names$Diagnosis %>% 
     filter(grepl('[Ff]racture', DESC)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)
#Smoker at the time of the fracture
Smoker <- tibble_names$History %>% 
     filter(grepl('[Ss]moker', DESC) & grepl('[Cc]urrent', DESC)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)
#A diagnosis of alcohol use (ICD-10 codes containing F10 or corresponding ICD-9 codes (303 alcohol dependance, 291 alcohol induced mental disorder, 3050 alcohol abuse, K70 alcoholic liver disease), which encompasses alcohol abuse, alcohol dependence and alcohol use, unspecified) is used as a surrogate for "3 or more units of alcohol per day" as specified by FRAX
EToH <- tibble_names$Diagnosis %>% 
     filter(grepl('F10|3050|291|303|V11.3|K70', CODE)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)
#RA diagnosis of Rheumatoid Arthritis 
RA <- tibble_names$Diagnosis %>% 
     filter(grepl('M05|M06|714', CODE)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)
#According to FRAX this includes osteoporosis (M81, 7330) and diseases strongly associated with osteoporosis including type I (insulin dependent) diabetes (E10, 250), osteogenesis imperfecta in adults (Q78.0, 75651), untreated long-standing hyperthyroidism E05, 242)(these are the codes for hyperthyroidism irrespective of treatment), hypogonadism (E29.1, 2572 (male), E28.3, 25631, 25639 (female) including) premature menopause (<45 years), chronic malnutrition (intestinal malabsorption K90, 579, nutritional malabsorption E41-46, 261-263, 579), or malabsorption and chronic liver disease (571-573, K71-K76)
#An alternative to malabsorption could be to consider BMI <15
Osteoporosis <- tibble_names$Diagnosis %>% 
     filter(grepl('M81|7330|E10|250|Q78.0|75651|E05|242|E29.1|2572|E28.3|25631|25639|k90|579|E41|E42|E43|E44|E45|E46|261|262|263|579|571|572|573|K71|K72|K73|K74|K75|K76', CODE)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)
#FRAX considers parental hip fracture a risk. We have substituted parental osteoporosis for this risk factor because there is a searchable diagnostic code
Parent_Osteoporosis <- tibble_names$Diagnosis %>% 
     filter(grepl('Z82.62|V17.81', CODE)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)
#Current exposure to oral glucocorticoids or exposure to oral glucocorticoids for more than 3 months at a dose eqivalent to prednisolone > 5mg daily
Glucocorticoid <- tibble_names$Meds %>% 
     filter(MODIFIER == 'mg' &
              grepl('[Bb]eclomethasone|[Bb]etamethasone|[Bb]udesonide|[Cc]ortisone|[Dd]examethasone|[Hh]ydrocortisone|[Mm]ethylprednisolone|[Pp]rednisolone|[Pp]rednisone|[Tt]riamcinolone', DESC)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE)

```
The below loop will collect and distill information relevant to calculating FRAX score FOR PATIENTS WHO HAVE HAD A FRACTURE ONLY. Each cycle of this loop will collect data for 1 patient and will populate the tibble 'EM_Dates'
```{r loop}
#Create an empty tibble ('EM_Dates') with the below columns. It will later be populated by each cycle of the loop
EM_Dates <- tibble(
  ID = numeric(),
  DATE = mdy_hms(),
  CODE = character(),
  DESC = character(),
  HT.in = numeric(), 
  WT.in = numeric(),
  Birth_Date = mdy_hms(),
  ageatEM = numeric(),
  ageatdeath = numeric(),
  GENDER = character(), 
  Og_RACE = character(), 
  CRF = numeric(),
  #CRF is Combined Risk Factors
  BMI = numeric()
)
#Tibble where the risk factors will be compiled
CRF <- tibble(
  ID = numeric(),
  Fracture = character(),
  Smoker = character(),
  EToH = character(),
  RA = character(),
  Osteoporosis = character(),
  Parent_Osteoporosis = character(),
  Glucocorticoid_rf = character()
)
#Collect risk factors into above tibble. If there is no observation prior to DATE, it is assumed that the risk factor was not present. For example, if there is no prior fracture documented, it is assumed that there was no prior fracture

for (i in 1:nrow(ID_EM)){
  id <- as.double(ID_EM$ID[i])
loop_ID1 <- ID_EM %>% 
    filter(ID == id)
#Below are common variables that will be used throughout the loop
date <- loop_ID1$DATE[1]
CODE <- loop_ID1$CODE[1]
DESC <- loop_ID1$DESC[1]
#REMOVE THIS IF YOU ARE NOT LOOKING FOR HIP FRACTURES ONLY!!!
#####
#Create a tibble with one patients into
loop_ID <- x %>% 
    filter(ID == id)
#Select only the info prior to fracture
loop_ID_prefrac <- loop_ID %>% filter(DATE <= date)
#Height and weight closest to the time of fracture
HT <- (loop_ID %>% 
         filter(CODE == "HT") %>% 
         arrange(desc(DATE)) %>% 
         select(NUMERIC_VALUE)
       )[1,1]
WT <- (loop_ID_prefrac %>% 
         filter(CODE == "WT") %>% 
         arrange(desc(DATE)) %>% 
         select(NUMERIC_VALUE)
       )[1,1]
#If there is no height or weight prior to the fracture date, use the oldest measurement
HT <- as.double(ifelse(is.na(HT), ((
  loop_ID %>% 
    filter(CODE == "HT") %>% 
    arrange(desc(DATE)) %>% 
    select(NUMERIC_VALUE))[1,1]),
  HT))
WT <- as.double(ifelse (is.na(WT), ((
  loop_ID %>% 
    filter(CODE == "WT") %>% 
    arrange(desc(DATE)) %>% 
    select(NUMERIC_VALUE))[1,1]), 
  WT))
#Calculate BMI
BMI <- (703*as.double(WT)/as.double(HT)^2)
yID <- y %>% filter(ID == id)
#Calcuate age at time of fracture
Birth_Date <- (yID %>% select(BIRTH_DATE))[[1]]
ageatEM <- (date - Birth_Date) %>% as.double/365.25
ageatdeath <- as.double(yID %>% select(AGE_AT_DEATH))
#Select gender
Gender <- yID %>% select(GENDER)
#Select race
Race <- yID %>% select(RACE)
# alternative to 'filter(grepl())' is library(stringr), x %>% filter(str_detect(rowname, "^ICD-10 code")). '^' means starts with
Fracture_loop <- as.character(
  (filter(Fracture, id == ID & DATE <= date) %>%
  select(DESC))[1,2]
  )
#Smoking within the last 2 years
Smoker_loop <- as.character(
  (filter(Smoker, id == ID & DATE <= date & DATE >= date-years(2)) %>% 
     select(DESC))[1,2]
  )
#A diagnosis of alcohol use (ICD-10 codes containing F10 or corresponding ICD-9 codes (303 alcohol dependance, 291 alcohol induced mental disorder, 3050 alcohol abuse), which encompasses alcohol abuse, alcohol dependence and alcohol use, unspecified) is used as a surrogate for "3 or more units of alcohol per day" as specified by FRAX
EToH_loop <- as.character(
  (filter(EToH, id == ID & DATE <= date) %>%
     select(CODE))[1,2]
  )
#A diagnosis of Rheumatoid Arthritis 
RA_loop <- as.character(
  (filter(RA, id == ID & DATE <= date) %>%
     select(CODE))[1,2]
  )
#According to FRAX this includes osteoporosis (M81, 7330) and diseases strongly associated with osteoporosis including type I (insulin dependent) diabetes (E10, 250), osteogenesis imperfecta in adults (Q78.0, 75651), untreated long-standing hyperthyroidism E05, 242)(these are the codes for hyperthyroidism irrespective of treatment), hypogonadism (E29.1, 2572 (male), E28.3, 25631, 25639 (female) including) premature menopause (<45 years), chronic malnutrition (intestinal malabsorption K90, 579, nutritional malabsorption E41-46, 261-263), or malabsorption and chronic liver disease
#An alternative to malabsorption could be to consider BMI <15
Osteoporosis_loop <- as.character(
  (filter(Osteoporosis, id == ID & DATE <= date) %>% 
     select(CODE))[1,2]
  )
#FRAX considers parental hip fracture a risk. We have substituted parental osteoporosis for this risk factor because there is a searchable diagnostic code
Parent_Osteoporosis_loop <- as.character(
  (filter(Parent_Osteoporosis, id == ID) %>% 
     select(CODE))[1,2]
  )
#Current exposure to oral glucocorticoids or exposure to oral glucocorticoids for more than 3 months at a dose eqivalent to prednisolone > 5mg daily
Glucocorticoid_data <- (
  filter(Glucocorticoid, id == ID & DATE <= date)
  )
Glucocorticoid_rf <- ifelse(Glucocorticoid_data[1,1] == 'NA', 'NA', Glucocorticoid_data[1,1])
#Exclude glucocorticoids taken for less than 3 months
#Add each risk factor to the 'CRF' tibble
CRF <- CRF %>% add_row(
ID = id,
Fracture = Fracture_loop,
Smoker = Smoker_loop,
EToH = EToH_loop,
RA = RA_loop,
Osteoporosis = Osteoporosis_loop,
Parent_Osteoporosis = Parent_Osteoporosis_loop,
Glucocorticoid_rf = as.character(Glucocorticoid_rf)
)
#Change the name of the tibble to calculate CRF score below so as to preserve the CRF tibble for retrospective analysis
CRF_calc <- CRF %>% filter(ID == id) %>% select(!ID)
#Convert risk factors to 0 (not present) or 1 (present)
CRF_calc[!is.na(CRF_calc)] <- '1'
CRF_calc[is.na(CRF_calc)] <- '0'
#Calculate CRF
CRF_score <- sum(as.double(CRF_calc))
#Put all the data you collected into 'EM_Dates'
EM_Dates = EM_Dates %>% add_row(
ID = id,
DATE = date,
CODE = CODE,
DESC = DESC,
HT.in = as.double(HT), 
WT.in = as.double(WT), 
Birth_Date = Birth_Date,
ageatEM = as.double(ageatEM),
ageatdeath = as.double(ageatdeath),
GENDER = as.character(Gender), 
Og_RACE = as.character(Race), 
CRF = as.double(CRF_score),
BMI = as.double(BMI)
)
}
#Save the original EM_Dates because the below chunks will make changes to it
EM_Dates_Og <- EM_Dates
#add time from fracture to death via mutate()
EM_Dates <- EM_Dates %>% mutate(time_to_death = ageatdeath - ageatEM)
#remove anyone < 50 years old at the first fracture. This should not include anyone in the cohort, but will catch anyone extra
EM_Dates <- EM_Dates %>% filter(ageatEM >= 50)
```
Calculate FRAX score below
```{r FRAX calculator}
#add death date to EM_Dates and change the name of the tibble to 'px'
px <- y %>%
  select(ID, DEATH_DATE) %>%
  right_join(EM_Dates, by = ID)
#Round BMI down to the nearest 5
floorfive <- function(x) {floor(x/5)*5}
px <- px %>% mutate(
  BMI = ifelse((BMI<45)&(BMI>15),floorfive(BMI), 
               ifelse (BMI>45,45,15)
               )
  )
#Round age down to the nearest 5
px <- px %>% mutate (
  AGE = ifelse((AGE<90)&(AGE>50),floorfive(AGE), 
               ifelse (AGE<50,50,90)
               )
  )
#Change race to white if not white, black, hispanic or asian
#Original race will be maintained the the column 'Og_RACE'
px$Og_RACE <- as.character(px$Og_RACE)
Races <- c("W","B","H","A")
px <- px %>% mutate(RACE = ifelse(
  (Og_RACE %in% Races), Og_RACE, "W")
  )

#Join FRAX and px to calculate 10 year risk of major osteoporotic fracture (do you mean major hip fracture?)
ed <- px %>% 
  left_join(FRAX, by= c("GENDER","RACE","AGE","CRF","BMI"), copy=FALSE, keep=FALSE)%>%
  select(ID, Risk, Range_upper, Range_lower) %>% 
  arrange(ID) %>%
  na.omit()

EM_Dates_ed <- EM_Dates %>% left_join(ed, by = 'ID')

write.csv(EM_Dates_ed, "C:\\Users\\acriscit\\Documents\\EM_Dates_ed.csv")
```
#Descriptors
```{r descriptors}
percent_complete_cases <- (nrow(ed)/nrow(EM_Dates))*100
percent_missing_FRAX <- (nrow(ed)/nrow(EM_Dates_ed))*100
percent_missing_HT <- (nrow(is.na(EM_Dates_ed$HT.in))/nrow(EM_Dates_ed))*100
percent_missing_WT <- (nrow(is.na(EM_Dates_ed$WT.in))/nrow(EM_Dates_ed))*100

Flow <- tibble()

Flow <- Flow %>% add_row(
  percent_complete_cases = as.double(percent_complete_cases),
  number_excluded_frac = as.double(number_excluded_frac),
  number_excluded_hip = as.double(number_excluded_hip),
  percent_missing_FRAX = as.double(percent_missing_FRAX),
  percent_missing_WT = as.double(percent_missing_WT),
  percent_missing_HT = as.double(percent_missing_HT)
)
write.csv(Flow, "C:\\Users\\acriscit\\Documents\\Flow.csv")

#Remove any tibbles you will not need
rm(EM_Dates)
#EM_Dates_ed contains all the info from the table
#EM_Dates_Og contains the original, unaltered info from the FRAX loop
```
#The below lines will select patients with a history of fracture. The patient pull in i2b2 stipulated that patients not have a diagnosis of fracture prior to the index date (ED visit between 5/1/2013 and 5/31/2014). Importantly, patients may have a fracture after 5/31/2014. We want to capture these patients.

#It will also allow you to select patients without fracture from a population of patients with and without fractures.
```{r filtering}
#Select the ID of anyone with a non MOF diagnosis: ICD9/10 codes- S62 hand, S92, S99 foot, M88 Osteodeformans/Paget's disease of bone, S02 Face/skull, V trauma due to a collision, Y04 assault by bodily force, C79.5 metastatic disease to bone
non_MOF <-  tibble_names$Diagnosis %>%
  #filter(grepl('[Ff]racture', DESC)) %>%
  filter(grepl('S62|S92|S99|M88|S02|V|Y04|C79.5', CODE)) %>%
  group_by(ID) %>%
  arrange(DATE, .by_group = TRUE) %>% 
  distinct(ID, .keep_all = TRUE) %>%
  select(ID, DATE, CODE, DESC)
### Not doing anything with the above info
```
#select patients who subsequently had a MOF
```{r MOF}
#Select anyone with MOF: hip fracture M84.359, M84.459, M84.659, Vertebral: 805|806|733.13|S12|M48.50|M80.08, Humerus (distal and shaft): 812.01|812.23|733.11|S42.2|M84.429A, Forearm: 813.45|813.8|733.12|813.0|S52|M84.429A
MOF <- tibble_names$Diagnosis %>%
  #filter(grepl('[Ff]racture', DESC)) %>%
  filter(grepl('M84.359|M84.459|M84.659|820|S72.0|S72.90XA|S72.91XA|S72.92XA|805|806|733.13|S12|M48.50|M80.08|812.01|812.23|733.11|S42.2|M84.429A|813.45|813.8|733.12|813.0|S52|M84.429A', CODE)) %>%
  group_by(ID) %>%
  arrange(DATE, .by_group = TRUE) %>% 
  distinct(ID, .keep_all = TRUE) %>%
  select(ID, DATE, CODE, DESC)
MOF <- rename(MOF, fx_DATE = DATE)

#Join the date of fracture (after EM visit) to EM_Dates_ed
EM_Dates_ed$ID <- as.character(EM_Dates_ed$ID)
EM_Dates_ed <- EM_Dates_ed %>% left_join(MOF, by = 'ID')
#Create a column to see when the fracture occurred
EM_Dates_ed <- EM_Dates_ed %>% mutate(fx_time = difftime(fx_DATE, DATE,
         units = c("days")))
#Will be positive if after the ED visit
#Note that there should not be any fractures prior to 5/31/2014, if there are, i2b2 let patients slip through the collection criteria
```
#Must discern which patients have diabetes within this cohort, and when they were diagnosed
