---
title: "FRAX calculator for I2b2 results 1/10/21"
author: "Anthony Criscitiello and Ellen Quillen"
date: "1/10/2021"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "FRAX calculator for I2b2 results"
output: html_document
---
```{r setup}
library(tidyverse)
library(readxl)
library(lubridate)
library(tibble)
library(purrr)
```
^I find it helpful to set everything up at the beggining and you can also make some modifications to how this Rmd turns into an html. Here is cheatsheet on options for R markdown: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
##Introduction
The goal of this is to .. (anything in white will show up as standard text, anything in gray will be analyzed as script)
You can decide how many chunks you want to break your script into. For me this is generally determined by the amount of related information I want to include.
I put my comments below in upper case to try and make them easier for you to find.
```{r load FRAX ppr charts and patient data}
#load FRAX paper charts (10-year risk of major osteoporotic fracture based on BMI)
#FRAX <-  read_csv("E:/Dropbox/Students/Anthony/FRAX.csv", header = TRUE, stringsAsFactors = FALSE)
FRAX <-  read_csv("C:\\Users\\acriscit\\Documents\\FRAX.csv", col_types = "ccdddddd")
names(FRAX)<- c("GENDER", "RACE", "AGE","CRF","BMI","Risk", "Range_upper", "Range_lower")
#Load I2b2 data
#xl_data <- "E:/Dropbox/Students/Anthony/20201020180719000_acriscit.xlsx"
FRAX_data <- "C:\\Users\\acriscit\\Documents\\20201020180719000_acriscit.xlsx"

#Label each list element with it's associated tab from the excel sheet
tab_names <- excel_sheets(path = FRAX_data)
#Below is the slow line
FRAX_data <- map(tab_names, ~read_excel(path = FRAX_data, sheet = .x))
FRAX_data <- FRAX_data %>% set_names(tab_names)
```
Now that the data is loaded into R we will start to process it into a more readable ('tidy') form
```{r tidy_frax}

#Collect the tabs from 'xl_data' that contain info used to calculate FRAX score
tab_names2 <- tibble(
Row = (1:length(tab_names)),
tab_names = tab_names
)

Frax_tabs <- tab_names2 %>% 
    filter(
    grepl('Diagnoses', tab_names) | 
    grepl('Labs', tab_names) | 
    grepl('Meds', tab_names) | 
    grepl('Vitals', tab_names) | 
    grepl('Smoking', tab_names)
  )
FRAX_pts <- FRAX_data[c(Frax_tabs$Row)]

#rename columns and remove spaces from column names
col_names <- c(
  "ID", 
  "VISIT", 
  "DATE",
  "CODE",
  "DESC", 
  "MODIFIER", 
  "TEXT_VALUE", 
  "NUMERIC_VALUE", 
  "UNITS"
)
for (i in 1:length(FRAX_pts)){
  names(FRAX_pts[[i]]) <- col_names[1:ncol(FRAX_pts[[i]])]
}

#Tidy the data by condensing the 'list' into a 'tibble' using 'bind_rows'
#IS THIS JUST A SUBSET/FILTER CALL? IT DOESN'T SEEM LIKE YOU ARE BINDING ROWS
#BEWARE USING X AND Y OR SIMILARLY NON-DESCRIPT VARIABLES FOR OBJECTS YOU NEED TO USE OVER A LONG PERIOD. MORE DESCRIPTIVE NAMES HELP YOU REMEMBER WHAT YOU WERE DOING DOWN THE LINE
x <- bind_rows(FRAX_pts)
x$DATE <- mdy_hms(x$DATE, quiet = TRUE,
  tz = "America/New_York"
)

#Collect the tabs containing demographic info used to calculate FRAX score from 'FRAX_data'
info_tabs <- tab_names2 %>% filter(grepl('Patients', tab_names))
FRAX_info <- FRAX_data[c(info_tabs$Row)]
col_names <- c(
  "ID", 
  "AGE_NOW", 
  "AGE_AT_DEATH", 
  "GENDER", 
  "RACE", 
  "MIXED RACE", 
  "ETHNICITY"
)
for (i in 1:length(FRAX_info)){
  names(FRAX_info[[i]]) <- col_names[1:ncol(FRAX_info[[i]])]
}
y <- bind_rows(FRAX_info)
```
Now that the patient data is more readable, we will create a tibble that contains data needed to calculate a FRAX score for each patient. This will primarily be accomplished with the use of a loop.
```{r calc_frax}

#Create a tibble ('Fracture_Dates') with columns for the date of the most recent fracture for each patient and associated FRAX characteristics
Fracture_Dates <- tibble(
  ID = numeric(),
  DATE = mdy_hms(),
  CODE = character(),
  DESC = character(),
  HT.in = numeric(), 
  HTDaysBeforeFrac = numeric(), 
  WT.in = numeric(), 
  WTDaysBeforeFrac = numeric(), 
  AGE = numeric(), 
  GENDER = character(), 
  Og_RACE = character(), 
  CRF = numeric(),
  #CRF is Combined Risk Factors
  BMI = numeric(),
  Number_of_fractures= numeric(),
  Risk = numeric(),
  Range_upper = numeric(),
  Range_lower = numeric()
)

#Create a tibble with all the IDs of patients who have had a fracture
all_fracs <- x %>% 
  filter((grepl('M80', CODE))) %>% 
  filter((grepl('fracture', DESC))) %>% 
  select(ID)
all_IDs <- tibble(y['ID'])
ID_fracs <- right_join(all_IDs, all_fracs, by = NULL) %>% distinct()

#The below loop is will collect and distill information relevant to calculating FRAX score
#Each cycle of this loop will collect data for 1 patient
for (i in 1:nrow(ID_fracs)){
  id <- as.double(ID_fracs[i,])
#Filtering by ICD-10 code 'M80' (pathologic fracture) and diagnosis described as 'fracture' to select the first fracture a patient had
pt_fractures <- (x %>% 
    filter(ID == id) %>% 
    filter((grepl('M80', CODE))) %>% 
    filter((grepl('fracture', DESC))) %>% 
    arrange(DATE))[1,]
date <- pt_fractures$DATE[1]

#Number of fractures the patient has experienced since the first
Number_of_fractures <- nrow(pt_fractures)

#Height and weight closest to the time of fracture
HT <- (x %>% 
         filter(CODE == "HT", ID == id) %>% 
         arrange(desc(DATE)) %>% 
         select(NUMERIC_VALUE)
       )[1,1]
WT <- (x %>% 
         filter(CODE == "WT", ID == id, DATE <= date) %>% 
         arrange(desc(DATE)) %>% 
         select(NUMERIC_VALUE)
       )[1,1]

#If there is no height or weight prior to the fracture date, use the oldest measurement
HT <- as.double(ifelse(HT == 0, ((
  x %>% filter(CODE == "HT", ID == id) %>% 
    arrange(desc(DATE)) %>% 
    select(NUMERIC_VALUE))[1,1]),
  HT))
WT <- as.double(ifelse (WT == 0, ((
  x %>% filter(CODE == "WT", ID == id) %>% 
    arrange(desc(DATE)) %>% 
    select(NUMERIC_VALUE))[1,1]), 
  WT))

#Calculate BMI
BMI <- (703*as.double(WT)/as.double(HT)^2)

#Calcuate age at time of fracture
#Rewrite this using birth dates
ageatfracture <- as.double(y %>% filter(ID == id) %>% select(AGE_NOW)) - (as.double(year(today())) - as.double(year(Fracture_Dates$DATE[i])))
#try this: ageatfracture <- as.Date(Fracture_Dates$DATE[i]) - as.Date(Birth_Date)

#Select gender
Gender <- y %>% filter(ID == id) %>% select(GENDER)

#Select race
Race <- y %>% filter(ID == id) %>% select(RACE)

#Tibble where the risk factors will be compiled
CRF <- tibble(
  Fracture = character(),
  Smoker = character(),
  EToH = character(),
  RA = character(),
  Osteoporosis = character(),
  Parent_Osteoporosis = character(),
  Glucocorticoid = character()
)
#Collect risk factors into above tibble from  (Fracture  Parent fracture Smoking Glucocorticoids RA  Alcohol use Secodary osteoporosis)
#If there is no observation prior to DATE, it is assumed that the risk factor was not present. For example, if there is no prior fracture, it is assumed that there was no prior fracture
ID_data <- x %>% filter(ID == id & DATE < date)
#'[Ff]racture'
Fracture <- as.character(
  (ID_data %>% 
     filter(grepl('Fracture|fracture', DESC)) %>% select(DESC))[1,1])
Smoker <- as.character(
  (ID_data %>% 
     filter(grepl('Smoker|smoker', DESC)) %>% select(DESC))[1,1])
EToH <- as.character(
  (ID_data %>% 
     filter(grepl('Alcoholism|alcoholism', DESC)) %>% select(DESC))[1,1])
RA <- as.character(
  (ID_data %>% 
     filter(grepl('Rheumatoid|rheumatoid', DESC),grepl('M05|M06|714', CODE)) %>% select(DESC))[1,1])
Osteoporosis <- as.character(
  (ID_data %>% 
     filter(grepl('Osteoporosis|osteoporosis', DESC), grepl('M80|733', CODE)) %>% select(DESC))[1,1])
#This should be parent hip fracture
Parent_Osteoporosis <- as.character(
  (ID_data %>% 
     filter(grepl('Z82.62|V17.81', CODE)) %>% select(CODE))[1,1])
#Current exposure to oral glucocorticoids or exposure to oral glucocorticoids for more than 3 months at a dose eqivalent to prednisolone > 5mg daily
Glucocorticoid <- as.character(
  Glucocorticoid <- (
  (x %>% 
     filter(DATE > (date - months(3)) & UNITS == 'mg' & grepl('sone| solone', DESC) & grepl('[Ii]njection| [Tt]ablet| [Ii]vbp', DESC)
     )))[1,1])

#Put this table into use to ensure dose is adequate
#steriod_conversion_table <- tibble(
#  steriod = c("cortisone", "hydrocortisone", "Methylprednisone", "Prednisolone", "Prednisone", "Triamcinolone", "Betamethasone", "Dexamethasone"),
#  Dose = c(5, 4, 0.8, 1, 1, 0.8, 0.12, 0.15)
#)
#Divide by the above number (bc the reference is 5mg prednisolone)
#To parse this out use ifelse statements. If it is injection/iv count it (ifelse(filter(grepl('IV or injection'))), 1, 0
#separate this column into multiple columns using tidy data to pivot
#These lines are held because they slow down the loop

#Add each risk factor to the 'CRF' tibble
#Note that there is no ID in this tibble. This is because we have selected data specific to one patient in line 188. Furthermore, we are still within the loop. Each cycle of a loop collects data for one patient
CRF <- CRF %>% add_row(
Fracture = Fracture,
Smoker = Smoker,
EToH = EToH,
RA = RA,
Osteoporosis = Osteoporosis,
Parent_Osteoporosis = Parent_Osteoporosis,
Glucocorticoid = Glucocorticoid
)
#Convert risk factors to 0 (not present) or 1 (present)
CRF[!is.na(CRF)] <- '1'
CRF[is.na(CRF)] <- '0'
#Calculate CRF
CRF_score <- sum(as.double(CRF))
#Put all the data you collected into 'Fracture_Dates'
Fracture_Dates = Fracture_Dates %>% add_row(
ID = id,
DATE = date,
CODE = pt_fractures$CODE[1],
DESC = pt_fractures$DESC[1],
HT.in = as.double(HT), 
WT.in = as.double(WT), 
AGE = as.double(ageatfracture), 
GENDER = as.character(Gender), 
Og_RACE = as.character(Race), 
CRF = as.double(CRF_score),
BMI = as.double(BMI),
Number_of_fractures = as.double(Number_of_fractures),
)
}
#end of loop
```
The tibble 'Fracture_Dates' now contains a row for each patient and associated information necessary to calculate FRAX score. Below we will use the 'FRAX' tibble create from publicly available FRAX paper charts (https://www.sheffield.ac.uk/FRAX/charts.aspx) to calculate the FRAX score for each patient.

Discern whether the patients with fracture have diabetes.
What is the duration of disease prior to fracture?
Were they well controlled?
```{Diabetes calculator}
DM_dates <- tibble (
  ID = id,
Dx_DATE = date,
Duration = time_from_fracture
)

all_DM <- x %>% 
  filter((grepl('E11', CODE))) %>% 
  filter((grepl('diabetes', DESC))) %>% 
  select(ID)
all_IDs <- tibble(y['ID'])
ID_DM <- right_join(all_IDs, all_fracs, by = NULL) %>% distinct()

for(i in 1:nrow(ID_DM)) {
  id <- as.double(ID_DM[i,])
  
  pt_DM <- (x %>% 
     filter(ID == id) %>% 
     filter((grepl('E11|250.00', CODE))) %>% 
     filter((grepl('diabetes', DESC))) %>% 
      arrange(DATE))
  DM_Dx_date <- pt_DM$DATE[1]
  
  time_from_fracture <- as.Date(Fracture_dates %>% filter(ID == id) %>% select(DATE)) - as.Date(DM_Dx_date)

  DM_dates <- DM_dates %>% add_row(
   ID = id,
   Dx_DATE = DM_Dx_date,
   Duration = time_from_fracture
)
}
#If you have more time use blood glucose labs to determine level of control by HbA1c. Maybe average the number every year, then every 3 months the year before the fracture. Taking into account the number of measurements.

```
#Calculate FRAX score below
```{r calc_frax}
px <- Fracture_Dates

#Round BMI down to the nearest 5
floorfive <- function(x) {floor(x/5)*5}
px <- px %>% mutate(
  BMI = ifelse((BMI<45)&(BMI>15),floorfive(BMI), 
               ifelse (BMI>45,45,15)
               )
  )

#Round age down to the nearest 5
px <- px %>% mutate (
  AGE = ifelse((AGE<90)&(AGE>50),floorfive(AGE), 
               ifelse (AGE<50,50,90)
               )
  )

#Change race to white if not white, black, hispanic or asian
#Original race will be maintained the the column 'Og_RACE'
px$Og_RACE <- as.character(px$Og_RACE)
Races <- c("W","B","H","A")
px <- px %>% mutate(RACE = ifelse(
  (Og_RACE %in% Races), Og_RACE, "W")
  )

#Join FRAX and px to calculate 10 year risk of major osteoporotic fracture
fx <- px %>% left_join(FRAX, by= c("GENDER","RACE","AGE","CRF","BMI"), copy = FALSE, keep = FALSE)%>%
  select(ID, Risk, Range_upper, Range_lower) %>% 
  arrange(ID)
#Add FRAX score to 'Fracture_Dates'
Fracture_Dates = Fracture_Dates %>% add_row(
  Risk = fx$Risk,
  Range_upper = fx$Range_upper,
  Range_lower = fx$Range_lower
)
#or
write.csv(fx, "E:/Dropbox/Students/Anthony/fx.csv")
```
Other questions to ask by modifying this script
  Current question: Does FRAX estimate fracture risk for DM and non-DM patients equally? and Do FRAX scores differ between those with and without diabetes?
Others:
  Does duration of diabetes alter fracture risk?
  Does FRAX score change leading up to a fracture?



