---
title: "FRAX calculator for Diabetic patients with fracture"
author: "Anthony Criscitiello and Ellen Quillen"
date: "1/10/2021"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "FRAX calculator for I2b2 results"
output: html_document
---
```{r setup}
library(tidyverse)
library(readxl)
library(readr)
library(lubridate)
library(tibble)
library(purrr)
```
^I find it helpful to set everything up at the beggining and you can also make some modifications to how this Rmd turns into an html. Here is cheatsheet on options for R markdown: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
##Introduction
```{r load FRAX ppr charts and patient data}
#load FRAX paper charts (10-year risk of major osteoporotic fracture based on BMI)
FRAX <-  read_csv("C:\\Users\\acriscit\\Documents\\FRAX.csv", col_types = "ccdddddd")
names(FRAX)<- c("GENDER", "RACE", "AGE","CRF","BMI","Risk", "Range_upper", "Range_lower")

Data_Request_1501_dem_01_06_2021 <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_dem_01_06_2021.csv")
Data_Request_1501_dx_01_06_2021 <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_dx_01_06_2021.csv")
Data_Request_1501_fam_hx_01_06_2021 <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_fam_hx_01_06_2021.csv")
Data_Request_1501_img_01_06_2021 <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_img_01_06_2021.csv")
Data_Request_1501_Labs_01_07_2021 <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_Labs_01_07_2021.csv")
Data_Request_1501_meds_01_06_2021 <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_meds_01_06_2021.csv")
Data_Request_1501_vitals_01_11_2021 <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_vitals_01_11_2021.csv")

Demographics <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_dem_01_06_2021.csv")[1:100,]
Diagnoses <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_dx_01_06_2021.csv")[1:100,]
History <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_fam_hx_01_06_2021.csv")[1:100,]
Imaging <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_img_01_06_2021.csv")[1:100,]
Labs <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_Labs_01_07_2021.csv")[1:100,]
Meds <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_meds_01_06_2021.csv")[1:100,]
Vitals <- read_csv("~/Data_Request_1501_01_06_2021/Data_Request_1501_vitals_01_11_2021.csv")[1:100,]


FRAX_data <- "C:\\Users\\acriscit\\Documents\\practice 20210109112736029_acriscit.xlsx"

#Label each list element with it's associated tab from the excel sheet
tab_names <- excel_sheets(path = FRAX_data)
#Below is the slow line
FRAX_data <- map(tab_names, ~read_excel(path = FRAX_data, sheet = .x))
FRAX_data <- FRAX_data %>% set_names(tab_names)

Patients <- FRAX_data$Patients
Diagnoses <- FRAX_data$Diagnoses
Labs <- FRAX_data$Labs
Meds <- FRAX_data$Meds
Vitals <- FRAX_data$Vitals
Smoking <- FRAX_data$Smoking
```
This data comes in multipe sheets unlike the I2b2 data which comes in 1 sheet with multiple tabs.

Now that the data is loaded into R we will start to process it into a more readable ('tidy') form
```{r tidy_frax}
tibble_names <- list(
  Vitals, 
  Meds, 
  Labs, 
  #Imaging, 
  Diagnoses,
  History
  )
tibble_titles <- c(
  "Vitals", 
  "Meds", 
  "Labs", 
  #"Imaging", 
  "Diagnoses",
  "History"
  )
names(tibble_names) <- tibble_titles
#rename columns and remove spaces from column names
col_names <- c(
  "ID", 
  "VISIT", 
  "DATE",
  "CODE",
  "DESC", 
  "TEXT_VALUE", 
  "NUMERIC_VALUE", 
  "UNITS",
  "MODIFIER"
)
#This loop applies the new column names to each list element in tibble names and converts dates to mdy_hms
for (i in 1:length(tibble_names)){
  names(tibble_names[[i]]) <- col_names[1:ncol(tibble_names[[i]])]
  tibble_names[[i]]$DATE <- mdy_hm(
    tibble_names[[i]]$DATE, 
    quiet = TRUE
    )
}
#Flatten list into a tibble and organize it so that information is grouped by ID and arranged by date
x <- tibble_names %>%
  bind_rows() %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE) %>% 
  distinct(ID, .keep_all = TRUE)

col_names <- c(
  "ID", 
  "AGE_NOW", 
  "AGE_AT_DEATH",
  "BIRTH_DATE",
  "DEATH_DATE",
  "MATCH",
  "GENDER", 
  "RACE", 
  "MIXED RACE", 
  "ETHNICITY"
)
names(Demographics) <- col_names[1:ncol(Demographics)]

Demographics$BIRTH_DATE <- mdy_hm(Demographics$BIRTH_DATE, quiet = TRUE)
#y <- Demographics

```
Now that the patient data is more readable, we will create a tibble that contains data needed to calculate a FRAX score for each patient. This will primarily be accomplished with the use of a loop.
```{r loop preparation}
#Create a tibble ('Fracture_Dates') with columns for the date of the most recent fracture for each patient and associated FRAX score variables
Fracture_Dates <- tibble(
  ID = numeric(),
  DATE = mdy_hms(),
  CODE = character(),
  DESC = character(),
  HT.in = numeric(), 
  WT.in = numeric(),
  AGE = numeric(), 
  GENDER = character(), 
  Og_RACE = character(), 
  CRF = numeric(),
  #CRF is Combined Risk Factors
  BMI = numeric(),
  first_visit_date = mdy_hms(),
  last_visit_date = mdy_hms(),
  num_visits_before_frac = numeric(),
  num_visits_after_frac = numeric(),
  osteo_med_start_date = mdy_hms(),
  osteo_med_start_dose = numeric(),
  osteo_med_last_date = mdy_hms(),
  osteo_med_name = as.character()
)

#The below lines will select patients with fracture then remove instances with traumatic fracture or fractures of the hands, feet, face or skull. It is assumed that all other fractures in this >50 y/o population are major osteoporotic fractures. Like the below lines it will then group fracture events by patient ID and arrange them by date, allowing us to select the first fracture if there are multiple using 'distinct ()'
ID_fracs <- x %>%
  filter(grepl('[Ff]racture', DESC)) %>%
  #Exclude S62 hand, S92, S99 foot, M88 Osteodeformans/Paget's disease of bone, S02 Face/skull, V trauma due to a collision, Y04 assault by bodily force
  filter(!grepl('S62|S92|S99|M88|S02|V|Y04', CODE)) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE) %>% 
  distinct(ID, .keep_all = TRUE)

#The below line will select patients with hip fracture
Hip_fracs <- x %>%
  #filter for ICD code M84.x59 which indicates a hip fracture where x = 3,4,5 or 6
  filter((grepl('M84&59', CODE))) %>%
  group_by(ID) %>% 
  arrange(DATE, .by_group = TRUE) %>% 
  distinct(ID, .keep_all = TRUE)
```

```{r preloop}
#Collect risk factors into above tibble. If there is no observation prior to DATE, it is assumed that the risk factor was not present. For example, if there is no prior fracture documented, it is assumed that there was no prior fracture

# alternative to 'filter(grepl())' is library(stringr), x %>% filter(str_detect(rowname, "^ICD-10 code")). '^' means starts with

Fracture <- x %>% 
     filter(grepl('[Ff]racture', DESC))

#Any smoking history
Smoker <- x %>% 
     filter(grepl('[Ss]moker', DESC) & !grepl('Never', DESC))
#A diagnosis of alcohol use (ICD-10 codes containing F10 or corresponding ICD-9 codes (303 alcohol dependance, 291 alcohol induced mental disorder, 3050 alcohol abuse), which encompasses alcohol abuse, alcohol dependence and alcohol use, unspecified) is used as a surrogate for "3 or more units of alcohol per day" as specified by FRAX
EToH <- x %>% 
     filter(grepl('F10|3050|291|303', CODE))
#A diagnosis of Rheumatoid Arthritis 
RA <- x %>% 
     filter(grepl('M05|M06|714', CODE))
#According to FRAX this includes osteoporosis (M81, 7330) and diseases strongly associated with osteoporosis including type I (insulin dependent) diabetes (E10, 250), osteogenesis imperfecta in adults (Q78.0, 75651), untreated long-standing hyperthyroidism E05, 242)(these are the codes for hyperthyroidism irrespective of treatment), hypogonadism (E29.1, 2572 (male), E28.3, 25631, 25639 (female) including) premature menopause (<45 years), chronic malnutrition (intestinal malabsorption K90, 579, nutritional malabsorption E41-46, 261-263), or malabsorption and chronic liver disease
#An alternative to malabsorption could be to consider BMI <15
Osteoporosis <- x %>% 
     filter(grepl('M81|7330|E10|250|Q78.0|75651|E05|242|E29.1|2572|E28.3|25631|25639|k90|579|E41|E42|E43|E44|E45|E46|261|262|263', CODE))
#FRAX considers parental hip fracture a risk. We have substituted parental osteoporosis for this risk factor because there is a searchable diagnostic code
Parent_Osteoporosis <- x %>% 
     filter(grepl('Z82.62|V17.81', CODE))
#Current exposure to oral glucocorticoids or exposure to oral glucocorticoids for more than 3 months at a dose eqivalent to prednisolone > 5mg daily
Glucocorticoid <- x %>% 
     filter(UNITS == 'mg' &
              grepl('[Bb]eclomethasone|[Bb]etamethasone|[Bb]udesonide|[Cc]ortisone|[Dd]examethasone|[Hh]ydrocortisone|[Mm]ethylprednisolone|[Pp]rednisolone|[Pp]rednisone|[Tt]riamcinolone', DESC) &
              grepl('[Ii]njection| [Tt]ablet| [Ii]vbp', DESC)
     )

#FDA approved osteoporosis medication usage
osteo_med_all <- x %>% 
  filter(
  grepl('[Ee]tidronate|[Aa]lendronate|[Ii]bandronate|[Zz]oledronic|[Rr]isedronate|[Cc]alcitonin|[Dd]enosumab', DESC)
  )

```
The below loop is will collect and distill information relevant to calculating FRAX score FOR PATIENTS WHO HAVE HAD A FRACTURE ONLY. Each cycle of this loop will collect data for 1 patient
```{r frax_loop}
for (i in 1:nrow(ID_fracs)){
  id <- as.double(ID_fracs$ID[i])

x_ID <- ID_fracs %>% 
    filter(ID == id)
#Below are common vairables that will be used throughout the loop
date <- x_ID$DATE[1]
CODE <- x_ID$CODE[1]
DESC <- x_ID$DESC[1]
loop_ID <- x %>% 
    filter(ID == id)
loop_ID_prefrac <- loop_ID %>% filter(DATE <= date)

#Information about number of visits and date of visits relative to fracture date
visits <- loop_ID %>% 
    arrange(DATE) %>% 
  distinct(DATE, .keep_all = TRUE)

first_visit_date <- visits$DATE[1]
last_visit_date <- (
  visits %>% 
    arrange(desc(DATE))
  )$DATE[1]
num_visits_before_frac <- visits %>% 
  filter(DATE < date) %>% 
  nrow()
num_visits_after_frac <- visits %>% 
  filter(DATE > date) %>% 
  nrow()


#Height and weight closest to the time of fracture
HT <- (loop_ID %>% 
         filter(CODE == "HT") %>% 
         arrange(desc(DATE)) %>% 
         select(NUMERIC_VALUE)
       )[1,1]
WT <- (loop_ID_prefrac %>% 
         filter(CODE == "WT") %>% 
         arrange(desc(DATE)) %>% 
         select(NUMERIC_VALUE)
       )[1,1]

#If there is no height or weight prior to the fracture date, use the oldest measurement
HT <- as.double(ifelse(is.na(HT), ((
  loop_ID %>% 
    filter(CODE == "HT") %>% 
    arrange(desc(DATE)) %>% 
    select(NUMERIC_VALUE))[1,1]),
  HT))
WT <- as.double(ifelse (is.na(WT), ((
  loop_ID %>% 
    filter(CODE == "WT") %>% 
    arrange(desc(DATE)) %>% 
    select(NUMERIC_VALUE))[1,1]), 
  WT))

#Calculate BMI
BMI <- (703*as.double(WT)/as.double(HT)^2)

yID <- y %>% filter(ID == id)
#Calcuate age at time of fracture
#ageatfracture <- as.double.difftime(date - (yID %>% select(BIRTH_DATE))[[1]])/365.25
ageatfracture <- as.double.difftime(as.double(yID %>% select(AGE_NOW))[[1]] - (2021 - as.double(year(date))))

#Select gender
#Change 'MATCH' back to 'GENDER'
Gender <- yID %>% select(MATCH)

#Select race
#Change 'GENDER' back to 'RACE'
Race <- yID %>% select(GENDER)

#Tibble where the risk factors will be compiled
CRF <- tibble(
  Fracture = character(),
  Smoker = character(),
  EToH = character(),
  RA = character(),
  Osteoporosis = character(),
  Parent_Osteoporosis = character(),
  Glucocorticoid_rf = character()
)
#Collect risk factors into above tibble. If there is no observation prior to DATE, it is assumed that the risk factor was not present. For example, if there is no prior fracture documented, it is assumed that there was no prior fracture

# alternative to 'filter(grepl())' is library(stringr), x %>% filter(str_detect(rowname, "^ICD-10 code")). '^' means starts with

Fracture_loop <- as.character(
  (left_join(x_ID, Fracture, by = "ID") %>%
  select(DESC))[1,1]
  )
#Any smoking history
Smoker_loop <- as.character(
  (left_join(x_ID, Smoker, by = "ID") %>% 
     select(DESC))[1,1]
  )
#A diagnosis of alcohol use (ICD-10 codes containing F10 or corresponding ICD-9 codes (303 alcohol dependance, 291 alcohol induced mental disorder, 3050 alcohol abuse), which encompasses alcohol abuse, alcohol dependence and alcohol use, unspecified) is used as a surrogate for "3 or more units of alcohol per day" as specified by FRAX
EToH_loop <- as.character(
  (left_join(x_ID, EToH, by = "ID") %>%
     select(CODE))[1,1]
  )
#A diagnosis of Rheumatoid Arthritis 
RA_loop <- as.character(
  (left_join(x_ID, RA, by = "ID") %>%
     select(CODE))[1,1]
  )
#According to FRAX this includes osteoporosis (M81, 7330) and diseases strongly associated with osteoporosis including type I (insulin dependent) diabetes (E10, 250), osteogenesis imperfecta in adults (Q78.0, 75651), untreated long-standing hyperthyroidism E05, 242)(these are the codes for hyperthyroidism irrespective of treatment), hypogonadism (E29.1, 2572 (male), E28.3, 25631, 25639 (female) including) premature menopause (<45 years), chronic malnutrition (intestinal malabsorption K90, 579, nutritional malabsorption E41-46, 261-263), or malabsorption and chronic liver disease
#An alternative to malabsorption could be to consider BMI <15
Osteoporosis_loop <- as.character(
  (left_join(x_ID, Osteoporosis, by = "ID") %>% 
     select(CODE))[1,1]
  )
#FRAX considers parental hip fracture a risk. We have substituted parental osteoporosis for this risk factor because there is a searchable diagnostic code
Parent_Osteoporosis_loop <- as.character(
  (left_join(x_ID, Parent_Osteoporosis, by = "ID") %>% 
     select(CODE))[1,1]
  )
#Current exposure to oral glucocorticoids or exposure to oral glucocorticoids for more than 3 months at a dose eqivalent to prednisolone > 5mg daily
Glucocorticoid_data <- left_join(x_ID, Glucocorticoid, by = "ID")
#Exclude glucocorticoids taken for less than 3 months
Glucocorticoid_start_date <- (Glucocorticoid_data %>% arrange(DATE) %>% select(DATE))[[1,1]]
Glucocorticoid_stop_date <- (Glucocorticoid_data %>% arrange(desc(DATE)) %>% select(DATE))[[1,1]]
Glucocorticoid_duration <- (Glucocorticoid_stop_date - Glucocorticoid_start_date)
Glucocorticoid_rf <- ifelse(Glucocorticoid_duration > days(90), Glucocorticoid_duration, "NA")

#Add each risk factor to the 'CRF' tibble
CRF <- CRF %>% add_row(
Fracture = Fracture,
Smoker = Smoker,
EToH = EToH,
RA = RA,
Osteoporosis = Osteoporosis,
Parent_Osteoporosis = Parent_Osteoporosis,
Glucocorticoid_rf = Glucocorticoid_rf
)
#Change the name of the tibble to calculate CRF score below so as to preserve the CRF tibble for retrospective analysis
CRF_calc <- CRF
#Convert risk factors to 0 (not present) or 1 (present)
CRF_calc[!is.na(CRF)] <- '1'
CRF_clac[is.na(CRF)] <- '0'
#Calculate CRF
CRF_score <- sum(as.double(CRF_calc))

#FDA approved osteoporosis medication usage
osteo_med <- left_join(x_ID, osteo_med_all, by = "ID") %>% 
                           arrange(DATE)
osteo_med_name <- as.character(osteo_med$DESC[1])
osteo_med_start_date <- (osteo_med %>% 
                           select(DATE))[[1,1]]
osteo_med_start_dose <- (osteo_med %>% 
                           select(NUMERIC_VALUE))[[1,1]]
osteo_med_last_date <- (osteo_med %>% 
                           arrange(desc(DATE)) %>% 
                           select(DATE))[[1,1]]

#Put all the data you collected into 'Fracture_Dates'
Fracture_Dates = Fracture_Dates %>% add_row(
ID = id,
DATE = date,
CODE = CODE,
DESC = DESC,
HT.in = as.double(HT), 
WT.in = as.double(WT), 
AGE = as.double(ageatfracture), 
GENDER = as.character(Gender), 
Og_RACE = as.character(Race), 
CRF = as.double(CRF_score),
BMI = as.double(BMI),
first_visit_date = first_visit_date,
last_visit_date = last_visit_date,
num_visits_before_frac = num_visits_before_frac,
num_visits_after_frac = num_visits_after_frac,
osteo_med_start_date = osteo_med_start_date,
osteo_med_start_dose = osteo_med_start_dose,
osteo_med_last_date = osteo_med_last_date,
osteo_med_name = osteo_med_name
)
}



